---
title: "Carbonate and alkalinity dynamics of fjord sediments"
subtitle: "Changing boundary conditions"
output: html_document
author: "Emil De Borger, Sebastiaan van de Velde, Ulrike Braeckman, Karline Soetaert, Astrid Hylén"
date: "2026-04-14"
---

# 1. Introduction

This script contains simulations of alkalinity dynamics in fjords with changing
boundary conditions for chapter X in the AGU monograph "Fjord biogeochemistry" 
edited by Laura Wehrmann, Mark J. Hopwood, and Elin M. K. D. Chiche.

We use the FEMSDIA model (Soetaert et al., 2026), an expansion upon the OMEXDIA
model of early diagenesis (Soetaert et al., 1996) and subsequent FESDIA model 
(Nmor et al., 2021). FEMSDIA models the dynamics of O2, C, N, P, S, Fe, Mg, carbonite
and aragonite through mineralization processes and (re)oxidation reactions, 
and tracks alkalinity changes associated to said processes. 
A full description of the model and model code can 
be found on XX.

In this study, we set up the FEMSDIA model to simulate what we considered as average
fjord sediments, with conditions typical for (sub-)Arctic fjords. This entails generally
high porosity of 0.75, bottom water temperature of 2 degrees C, and bottom water 
conditions that generally resemble those of inflowing oceanic waters. On these 
base conditions, we test the importance of different boundary conditions that are
expected to play a strong role in governing carbonate and alkalinity dynamics and
range broadly between fjords and fjord systems globally: the sediment mass accumulation 
rate, the organic matter deposition rate, the reactivity of the organic matter,
the bottom water oxygen levels, the accumulation of iron, and the activities of 
fauna on the seafloor through bioturbation and bioirrigation.

Investigating the effect of deposition on alkalinity dynamics is not straight-
forward, as this encompasses several aspects. Deposition is the combination 
of mineral deposits and organic matter. Mineral deposits, vary seasonally as they
are related to sources from land, in this case glacial river runoff, and snowmelt 
and precipitation on the slopes surrounding fjords. Depending on the geology of 
the fjord catchment, specific minerals relevant to biogeochemical processes, such 
as iron rich materials, are associated with this mineral deposit. In the model, 
mineral deposits contribute to the sediment advection rate "w", which affects both
solids and solutes, and the deposition fluxes of reactive iron (FeOH3), and 
manganese(MnO2) to a lesser degree.
Organic material deposition also varies, this is the result of algae blooms, both 
within and outside of the fjord, of which a fraction ends up on the seafloor. Organic 
material also comes from the land surrounding the fjords, a combination of terrestrial
plants and plant remains, but also human inputs such as waste, or fish farming 
depositions. In the model these inputs are governed by the "Cflux" parameter, the 
overall input of (reactive) organic matter, of which the "quality" - i.e. the freshness,
the reactivity, can be governed with additional parameters.
Whereas these parameters can often be strongly correlated, they do not always have
to be. As fjord environments are enormously diverse, no general rules are available
to constrain several of these parameters simultaneously. So below, we vary these
boundary conditions individually, and assess their effect on alkalinity generation
in the seafloor. These are simplistic assessments, but they off insight in general 
dynamics.


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
library(FEMSDIA)
library(ggplot2)
library(patchwork)
library(tidyr)
library(dplyr)
library(purrr)
library(ReacTran)
library(ggh4x)
source("plotfuncs.R")

install.packages("gitcreds")
library(gitcreds)

Sys.which("git")
gitcreds_set()

devtools::install_github("DiagenesisHub/FEMSDIA", force = TRUE)

```

# 2. Background parametrization

This set of parameters was chosed based on Berg et al. 2003 (DOI: 10.2475/ajs.303.10.905),
whose model is similar, but not completely the same, which is why not all parameter
values are the same and tweaking was necessary to get similar profiles.

```{r}

baseParms <- list(Cflux         = 4.5         ,  # Initial guess - global fjord average (https://doi.org/10.1007/s12237-025-01582-5)
                  pFast         = 0.25         ,  # Berg et al. - split of 75slow/25 fast after 37 days.
                  rFast         = 75/365.25   ,  
                  rSlow         = 0.14/365.25  ,  
                  NCrFdet       = 12/106      ,  # Slightly more degraded OM in mixture
                  NCrSdet       = 12/106      ,  # Slightly more degraded OM in mixture
                  NCrRdet       = 12/106      ,  # Slightly more degraded OM in mixture
                  FeOH3flux     = 0.5         ,  
                  MnO2flux      = 14/365.25   ,  
                  O2bw          = 280         , # Reasonable estimate towards well oxygenated conditions (Watts et al., 2024 https://doi.org/10.1029/2023JG007978)
                  w             = 0.005/365.25,  # Range: 0.1 - 12 cm/yr (Bianchi et al., 2020 table 3: https://doi.org/10.1016/j.earscirev.2020.103145)
                  por0          = 0.75        ,  # Depositional environment 
                  pordeep       = 0.7         ,  # Depositional environment
                  rnitri1       = 5           ,  
                  rnitri2       = 5           ,  
                  Twater        = 2           ,
                  biot          = 0.001/365.25,
                  biotdepth     = 0.06        ,
                  DBL           = 0.0005      , 
                  irr           = 12/365.25   ,  # based on Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905)
                  irrdepth      = 0.03        ,
                  irratt        = 0.02        ,
                  ALKbw         = 2000        ,
                  ksO2oxic      = 3           ,
                  kinO2anox     = 1           ,
                  ksFeOH3       = 1.e+04      , 
                  kinFeOH3      = 1.e+04      , 
                  rH2SFeox      = 1.4*10^-9 * 86400,                 
                  ksSO4BSR      = 2e+03       , 
                  kinSO4Met     = 2e+03       ,       
                  rH2Sox        = 4.38*10^-6  , #,# Wang and Van Capellen 1996
                  rMnox         = 0.0137      , # Van Cappellen and Wang 1996
                  rFeox         = 0.294      ,             # Van Cappellen and Wang 1996
                  rFeS2prod_S0  = 2.3*10^-10 * 86400 /50, # Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905) # /50 otherwise way too much FeS2
                  rFeS2prod_HS  = 8.9*10^-10 * 86400 /50, # Rickard, 1997a # /50 otherwise way too much FeS2
                  rFeMnox       = 7.5*10^-9 * 86400     , # Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905) orig: 7.5*10^-11
                  rFeSprod      = 4.7*10^-8 * 86400     , # Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905) -7 expo original
                  rH2SMnox      = 2*10^-9 * 86400       , 
                  rFeSox        = 6.0*10^-7 * 86400     ,
                  rFeS2ox       = 3.0*10^-11 * 86400      
                  )

```

# 3. Sensitivity analyses
## 3.1. Sediment advection rate

Measured advection rates in fjords globally range from 0.01 cm y-1 to 12 cm y-1 
(Bianchi et al., 2020 - https://doi.org/10.1016/j.earscirev.2020.103145), though 
cases of extreme deposition have been observed of up to 13 m y-1 proximal to Muir
Glacier within 1 km of the grounding line (https://doi.org/10.1130/SPE261-p61).
This range stems from global differences, but also from within-fjord differences related to 
proximity to the glacier head. w tends to be highest closes to the head (ice-proximal)
as this region recieves input directly, of which a significant fraction sinks 
immediately. For this run the model domain needed to be extended down to extends 
to 500 m depth, the depth needed for highest advection rates to resolve and satisfy
the zero-flux boundary.

### 3.1.1. Runs

```{r advectionvariation}

# Vector of input w's - range containing results of Bianchi et al., (2020) extended
# towards extremely high advection rates near glacier termini.

inputw_cm_yr <- c(0.01, 0.05, 0.1, 0.2, 1, 5, 10, 50, 100, 400, 600, 1000)
inputw_m_day <- inputw_cm_yr / 100 / 365.25

# Define grid deep enough for zero gradient boundary
deepGrid <- setup.grid.1D(x.up = 0, dx.1 = 0.0001, N = 100, L = 500)
  
# save each wrun
runs_w <- vector("list", length(inputw_m_day))

for (i in seq_along(inputw_m_day)) {
  
  parms_i <- baseParms
  parms_i$w <- inputw_m_day[i]
  
  message(
    "Running w = ", inputw_cm_yr[i], " cm yr^-1",
    " = ", signif(inputw_m_day[i], 4), " m d^-1"
  )
  
  

  out_i <- FEMSDIA_run_steady(parms = parms_i, 
                              dynamicpH = TRUE,
                              runsteady_times = c(0, 1e+14),
                              Grid = deepGrid)
  
  runs_w[[i]] <- list(
    w_cm_yr = inputw_cm_yr[i],
    w_m_day = inputw_m_day[i],
    parms   = parms_i,
    output  = out_i
  )
}

# useful names
names(runs_w) <- paste0("w_", inputw_cm_yr, "_cm_yr")

# save all model runs
saveRDS(runs_w, file = "FEMSDIA_w_sensitivity_runs.rds")

runs_w <- readRDS("FEMSDIA_w_sensitivity_runs.rds")

```

### 3.1.2. Output generation

#### Plot DIC - Alkalinity - Organic C in depth for increasing w

```{r, fig.height = 3, fig.width = 8}

w_plotDAO <- plotDAO(unitplot = "(cm~yr^{-1})",
                        breaks = c(0.01, 0.1, 1, 10),
                        labels = as.character(c(0.01, 0.1, 1, 10)),
                        runs = runs_w,
                        varia = "w~")

w_plotDAO[[2]]

```

#### Budgets

##### C mineralization per mineralization process

```{r}

w_plotCbudg <- plotCbudg(runs = runs_w,
                          unitplot = "(cm~yr^{-1})",
                          varia = "w~")

w_plotCbudg[[2]]
```

##### DIC production / burial inorganic

```{r}

w_DICinorg <- plotDICinorg(runs = runs_w,
                           unitplot = "(cm~yr^{-1})",
                           varia = "w~")
w_DICinorg[2]

```

##### Alkalinity production per mineralization process - Krumins

```{r}

w_plotAlkProd <- plotAlkProd(runs = runs_w,
                             unitplot = "(cm~yr^{-1})",
                             varia = "w~")
w_plotAlkProd[2]

```

##### Alkalinity budget -Krumins analogue

```{r}

## Alkalinity budget - Krumins analogue

w_plotAlkKrumins <- plot_AlkKrumins(runs = runs_w,
                                    unitplot = "(cm~yr^{-1})",
                                    varia = "w~")
w_plotAlkKrumins[[2]]

```

##### Alkalinity budget - onlylongterm

```{r, fig.height = 4, fig.width = 7}

w_plotAlkLongterm <- plot_netalklongtermMAR(runs = runs_w,
                                    unitplot = expression("g m"^{-2}~"y"^{-1}),
                                    varia = "Mass~accumulation~rate~")
w_plotAlkLongterm[[2]]

ggsave("./outputfigures/advectionalk.pdf",
       height = 4, width = 8, units = "in",
       dpi = 300)

```

## 3.2. OM deposition flux - accumulation rate

The endpoints of the range of organic matter deposition values, 0.41 - 51.2 +- 4.3
mmol C m-2 d-1 are determined by Smith et al. (2015) on the low end and Walinsky 
et al.(2009) on the high end.

### 3.2.1. Runs

```{r label, options}

inputCmmol <- c(0.4, 1, 2, 5, 10, 15, 20, 25, 37.5, 50)

# save each wrun
runs_Cflux <- vector("list", length(inputCmmol))

for (i in seq_along(inputCmmol)) {
  
  parms_i <- baseParms
  parms_i$Cflux <- inputCmmol[i]
  
  message(
    "Running Cflux = ", inputCmmol[i], " mmol m^-2 yr^-1",
    " = ", signif(inputCmmol[i], 2), " mmol m^-2 yr^-1"
  )
  
  out_i <- FEMSDIA_run_steady(parms = parms_i, dynamicpH = TRUE,
                              runsteady_times = c(0, 1e+8)) 
  
  runs_Cflux[[i]] <- list(
    Cflux_m2day = inputCmmol[i],
    parms   = parms_i,
    output  = out_i
  )
}

# useful names
names(runs_Cflux) <- paste0("Cflux_", inputCmmol, "_mmol_/m2_/y")

# save all model runs
saveRDS(runs_Cflux, file = "FEMSDIA_Cflux_sensitivity_runs.rds")

runs_Cflux <- readRDS("FEMSDIA_Cflux_sensitivity_runs.rds")

```

### 3.2.2. Output generation
#### Plot DIC - Alkalinity - Organic C in depth for increasing C flux

```{r, fig.height = 3, fig.width = 8}

C_plotDAO <- plotDAO(unitplot = expression("mmol m"^{-2}~"d"^{-1}),
                        breaks = c(0.01, 0.1, 1, 10),
                        labels = as.character(c(0.01, 0.1, 1, 10)),
                        runs = runs_Cflux,
                        varia = "Cflux~")

C_plotDAO[[2]]

```

#### Budgets

##### C mineralization per mineralization process

```{r}

C_plotCbudg <- plotCbudg(runs = runs_Cflux,
                          unitplot = expression("mmol m"^{-2}~"d"^{-1}),
                          varia = "Cflux~")

C_plotCbudg[[2]]

```

##### DIC production / burial inorganic

```{r}

C_DICinorg <- plotDICinorg(runs = runs_Cflux,
                            unitplot = expression("mmol m"^{-2}~"d"^{-1}),
                            varia = "Cflux~")
C_DICinorg[2]

```

##### Alkalinity production per mineralization process - Krumins

```{r}

C_plotAlkProd <- plotAlkProd(runs = runs_Cflux,
                            unitplot = expression("mmol m"^{-2}~"d"^{-1}),
                            varia = "Cflux~")
C_plotAlkProd[2]

```

##### Alkalinity budget -Krumins analogue

```{r}

## Alkalinity budget - Krumins analogue
C_plotAlkKrumins <- plot_AlkKrumins(runs = runs_Cflux,
                            unitplot = expression("mmol m"^{-2}~"d"^{-1}),
                            varia = "Cflux~")
C_plotAlkKrumins[[2]]

```

##### Alkalinity budget - onlylongterm

```{r, fig.height = 4, fig.width = 7}

C_plotAlkLongterm <- plot_netalklongterm(runs = runs_Cflux,
                            unitplot = expression("mmol m"^{-2}~"d"^{-1}),
                            varia = "OC~deposition~flux~")
C_plotAlkLongterm[[2]]

ggsave("./outputfigures/cflux.pdf",
       height = 4, width = 8, units = "in",
       dpi = 300)

```

## 3.3. Iron deposition - accumulation rate

Iron is deposited at different rates depending on the weathering intensity and the 
geology of the fjord catchment. Wehrmann et al. (2014 - 10.1016/j.gca.2014.06.007)
report fluxes that range from 0.44 to 29.44 mmol m-2 d-1. However, evidence shows 
that 0.6 - 12 % of this inital flux is bioavailable-reactive (Laufer-Meiser et al., 
2021 - 10.1038/s41467-021-21558-w). Over time, though, these non-reactive strongly
crystalline forms may be broken down, and become biologically available and increase
the proportion of bioavailable iron forms up to ninefold. This is not a direct depo-
sition, but represents a source of reactive iron generated within the sediment itself.
Below, we test a range of reactive iron deposition fluxes of 0.026 - 3.533 mmol m-2 d-1,
corresponding to the 0.6 - 12 % range, as the our model does not include the initial
transformation of a non-reactive deposited iron fraction into bioavailable forms.
The highest deposition flux of reactive iron that could be modelled was 3.3 mmol m-2 d-1.
For higher values unrealistic bioirrigation values or process rates needed to be 
implemented, which we chose not to do. This may be a consequence of the simplistic
"only change one variable at the time" approach.

### 3.3.1. Runs

```{r label, options}

# Vector of reactive Fe deposition fluxes
inputFemmol <- c(0.025, 0.05, 0.1, 0.2, 0.4, 0.8, 1, 1.5, 2, 2.5, 3, 3.3)
deepGrid <- setup.grid.1D(x.up = 0, dx.1 = 0.0001, N = 100, L = 5)

runs_Fe <- vector("list", length(inputFemmol))

for (i in seq_along(inputFemmol)) {
  
  parms_i <- baseParms
  parms_i$FeOH3flux <- inputFemmol[i]
  
  message(
    "Running Fe = ", inputFemmol[i], "mmol m^-2 yr^-1"
  )
  
  out_i <- FEMSDIA_run_steady(parms = parms_i, 
                              dynamicpH = TRUE)
  
  runs_Fe[[i]] <- list(
    Feflux_m2day = inputFemmol[i],
    parms   = parms_i,
    output  = out_i
  )
}

# useful names
names(runs_Fe) <- paste0("Fe_", inputFemmol, "_mmolm2")

# save all model runs
saveRDS(runs_Fe, file = "FEMSDIA_Fe_sensitivity_runs.rds")

# Load
runs_Fe <- readRDS("FEMSDIA_Fe_sensitivity_runs.rds")

```

### 3.3.2. Output generation

#### Plot DIC - Alkalinity - Organic C in depth for increasing reactive Fe flux

```{r, fig.height = 3, fig.width = 8}

Fe_plotDAO <- plotDAO(unitplot = expression("mmol m"^{-2}~"d"^{-1}),
                        breaks = c(0.01, 0.1, 1, 3),
                        labels = as.character(c(0.01, 0.1, 1, 3)),
                        runs = runs_Fe,
                        varia = "Feflux~")

Fe_plotDAO[[2]]

```

#### Budgets

##### C mineralization per mineralization process

```{r}

Fe_plotCbudg <- plotCbudg(runs = runs_Fe,
                          unitplot = expression("mmol m"^{-2}~"d"^{-1}),
                          varia = "Feflux~")

Fe_plotCbudg[[2]]

```

##### DIC production / burial inorganic

```{r}

Fe_DICinorg <- plotDICinorg(runs = runs_Fe,
                            unitplot = expression("mmol m"^{-2}~"d"^{-1}),
                            varia = "Feflux~")
Fe_DICinorg[2]

```

##### Alkalinity production per mineralization process - Krumins

```{r}

Fe_plotAlkProd <- plotAlkProd(runs = runs_Fe,
                            unitplot = expression("mmol m"^{-2}~"d"^{-1}),
                            varia = "Feflux~")
Fe_plotAlkProd[2]

```

##### Alkalinity budget -Krumins analogue

```{r}
## Alkalinity budget - Krumins analogue
Fe_plotAlkKrumins <- plot_AlkKrumins(runs = runs_Fe,
                            unitplot = expression("mmol m"^{-2}~"d"^{-1}),
                            varia = "Feflux~")
Fe_plotAlkKrumins[[2]]

```
##### Alkalinity budget - onlylongterm

```{r}

Fe_plotAlkLongterm <- plot_netalklongterm(runs = runs_Fe,
                                    unitplot = expression("mmol m"^{-2}~"d"^{-1}),
                                    varia = "w~")
Fe_plotAlkLongterm[[2]]

ggsave("./outputfigures/fefluxalk.pdf",
       height = 4, width = 8, units = "in",
       dpi = 300)

```

## 3.4. the role of oxygenation

Exploring the range from anoxic bottom water to fully oxygenated (default).
These simulations start at 0.1 mmol O2 m-2 d-1 over true 0, as this state did not
solve.

### 3.4.1. Runs

```{r}

# Vector of input Oxygen fluxes
inputOxmmol <- c(0.1, 1, 2, 5, 10, 20, 50, 75, 100, 150, 200, 250, 300)

# save each wrun
runs_Ox <- vector("list", length(inputOxmmol))

for (i in seq_along(inputOxmmol)) {
  
  parms_i <- baseParms
  parms_i$O2bw <- inputOxmmol[i]
  
  message(
    "Running Ox = ", inputOxmmol[i], "mmol m^-3"
  )
  
  out_i <- FEMSDIA_run_steady(parms = parms_i, dynamicpH = TRUE)
  
  runs_Ox[[i]] <- list(
    Oxbw_m2 = inputOxmmol[i],
    parms   = parms_i,
    output  = out_i
  )
}

# useful names
names(runs_Ox) <- paste0("Ox_", inputOxmmol, "_mmolm3")

# save all model run
saveRDS(runs_Ox, file = "FEMSDIA_Ox_sensitivity_runs.rds")

runs_Ox <- readRDS("FEMSDIA_Ox_sensitivity_runs.rds")

```

### 3.4.2. Output generation

#### Plot DIC - Alkalinity - Organic C in depth for increasing reactive Fe flux

```{r, fig.height = 3, fig.width = 8}

Ox_plotDAO <- plotDAO(unitplot = expression("mmol m"^{-3}),
                        breaks = c(0.1, 1, 30, 100, 300),
                        labels = as.character(c(0.1, 1, 30, 100, 300)),
                        runs = runs_Ox,
                        varia = "OxBW ~")
Ox_plotDAO[[2]]

```

#### Budgets

##### C mineralization per mineralization process

```{r}

Ox_plotCbudg <- plotCbudg(runs = runs_Ox,
                          unitplot = expression("mmol m"^{-2}~"d"^{-1}),
                          varia = "OxBW~")

Ox_plotCbudg[[2]]

```

##### DIC production / burial inorganic

```{r}

Ox_DICinorg <- plotDICinorg(runs = runs_Ox,
                            unitplot = expression("mmol m"^{-3}),
                            varia = "OxBW~")
Ox_DICinorg[2]

```

##### Alkalinity production per mineralization process - Krumins

```{r}

Ox_plotAlkProd <- plotAlkProd(runs = runs_Ox,
                            unitplot = expression("mmol m"^{-3}),
                            varia = "OxBW~")
Ox_plotAlkProd[2]

```

##### Alkalinity budget -Krumins analogue

```{r}
## Alkalinity budget - Krumins analogue
Ox_plotAlkKrumins <- plot_AlkKrumins(runs = runs_Ox,
                            unitplot = expression("mmol m"^{-3}),
                            varia = "OxBW~")
Ox_plotAlkKrumins[[2]]

```

##### Alkalinity budget - onlylongterm

```{r}

Ox_plotAlkLongterm <- plot_netalklongterm(runs = runs_Ox,
                                    unitplot = expression("(mmol m"^{-3}~")"),
                                    varia = "Bottomwater~oxygen~")
Ox_plotAlkLongterm[[2]]


ggsave("./outputfigures/oxygenationalk.pdf",
       height = 4, width = 8, units = "in",
       dpi = 300)

```

## 3.5. The role of faunal activities

### 3.5.1. bioturbation

We apply a range of biodiffusion (Db - biot) rates that represent the range of
very low to intermediate bioturbation rates. High bioturbation is not expected in 
fjords due to colder temperatures and generally less labile organic carbon.
With increasing bioturbation also mixing depth is expected to scale (Boudreau, Middelburg)
according to the relation L = 4.6 sqrt(Db/k) with k the average reactivity of the
organic matter pool. This deepening bioturbation is applied simultaneously.

#### 3.5.1.1. Runs

```{r}

inputBiotcm2yr <- c(0.001, 0.005, 0.01, 0.05, 0.1, 1, 5, 10, 20, 50)
inputBiotm2d   <- inputBiotcm2yr / 365.25 / 10000
mixdepth       <- 4.6*sqrt(inputBiotcm2yr/22.9) /100

# save each wrun
runs_Biot <- vector("list", length(inputBiotm2d))

for (i in seq_along(inputBiotm2d)) {

  parms_i <- baseParms
  parms_i$biot <- inputBiotm2d[i]
  parms_i$biotdepth <- mixdepth[i]

  message(
    "Running Biot= ", inputBiotcm2yr[i], "cm^2 yr^-1",
    " = ", signif(inputBiotcm2yr[i], 2), " cm^2 yr^-1"
  )

  out_i <- FEMSDIA_run_steady(parms = parms_i,
                              dynamicpH = TRUE,
                              runsteady_times = c(0, 1e+8))

  runs_Biot[[i]] <- list(
    Biot_cm2d = inputBiotcm2yr[i],
    parms   = parms_i,
    output  = out_i
  )
}

# useful names
names(runs_Biot) <- paste0("Bioturbation_", inputBiotcm2yr, "_cm2d-1")
# save all model runs
saveRDS(runs_Biot, file = "FEMSDIA_Biot_sensitivity_runs.rds")
runs_Biot <- readRDS("FEMSDIA_Biot_sensitivity_runs.rds")

```

#### 3.5.1.2. Output generation

##### Plot DIC - Alkalinity - Organic C in depth for increasing reactive Fe flux

```{r, fig.height = 3, fig.width = 8}

Biot_plotDAO <- plotDAO(unitplot = expression("cm"^{2}~"y"^{-1}),
                        breaks = c(0.001, 0.1, 1, 10, 50),
                        labels = as.character(c(0.001, 0.1, 1, 10, 50)),
                        runs = runs_Biot,
                        varia = "Biodiffusion~coeff.~")

Biot_plotDAO[[2]]

```

##### Budgets

###### C mineralization per mineralization process

```{r}

Biot_plotCbudg <- plotCbudg(runs = runs_Biot,
                          unitplot = expression("cm"^{2}~"y"^{-1}),
                          varia = "Biodiffusion~coeff.~")

Biot_plotCbudg[[2]]

```

###### DIC production / burial inorganic

```{r}

Biot_DICinorg <- plotDICinorg(runs = runs_Biot,
                            unitplot = expression("cm"^{2}~"y"^{-1}),
                            varia = "Biodiffusion~coeff.~")
Biot_DICinorg[2]

```

###### Alkalinity production per mineralization process - Krumins

```{r}

Biot_plotAlkProd <- plotAlkProd(runs = runs_Biot,
                            unitplot = expression("cm"^{2}~"y"^{-1}),
                            varia = "Biodiffusion~coeff.~")
Biot_plotAlkProd[2]

```

###### Alkalinity budget -Krumins analogue

```{r}
## Alkalinity budget - Krumins analogue
Biot_plotAlkKrumins <- plot_AlkKrumins(runs = runs_Biot,
                            unitplot = expression("mmol m"^{-3}),
                            varia = "Biodiffusion~coeff.~")
Biot_plotAlkKrumins[[2]]

```

###### Alkalinity budget - onlylongterm

```{r}

Biot_plotAlkLongterm <- plot_netalklongterm(runs = runs_Biot,
                                    unitplot = expression("cm"^{2}~"y"^{-1}),
                                    varia = "Biodiffusion~coeff.~")
Biot_plotAlkLongterm[[2]]

ggsave("./outputfigures/bioturbationalk.pdf",
       height = 4, width = 8, units = "in",
       dpi = 300)

```

### 3.4.2. bioirrigation

Bioirrigation is a complex process, where faunal actions (feeding, respiration, movement)
cause an exchange of water between the overlying water column, and sediment porewater.
Here, we assume a standard bioirrigation rate that decreases exponentially  from 
sediment water interface to deeper sediment layers. Two analyses are performed, 
one where the attenuation coefficient is changed (how rapid does the bioirrigation
rate decrease/attenuate to 0 in depth), and the boirrigation rate itself (which
proportion of the porewater is exchanged per unit of time). both deliver 
different dynamics.

#### 3.4.2.1. First version with attenuation

```{r}

inputIrrAtt <- c(0.001, 0.005, 0.01, 0.02, 0.05, 0.1, 0.2, 0.5, 1, 2)

# save each wrun
runs_Irr <- vector("list", length(inputIrrAtt))

for (i in seq_along(inputIrrAtt)) {
  
  parms_i <- baseParms
  # parms_i$irr <- inputIrr[i]
  parms_i$irr <- 0.1
  parms_i$irratt <- inputIrrAtt[i]
  
  
  message(
    "Running Irr= ", inputIrrAtt[i], " m^-1",
    " = ", signif(inputIrrAtt[i], 2), " m^-1"
  )
  
  out_i <- FEMSDIA_run_steady(parms = parms_i, 
                              dynamicpH = TRUE,
                              runsteady_times = c(0, 1e14)
                              )
  
  runs_Irr[[i]] <- list(
    Irr_cm2d = inputIrrAtt[i],
    parms   = parms_i,
    output  = out_i
  )
}

# useful names
names(runs_Irr) <- paste0("Irrigation_", inputIrrAtt, "_m-1")

# save all model runs
saveRDS(runs_Irr, file = "FEMSDIA_Irr_sensitivity_runs.rds")

runs_Irr <- readRDS("FEMSDIA_Irr_sensitivity_runs.rds")

```

##### Output generation

###### Plot DIC - Alkalinity - Organic C in depth for increasing reactive Fe flux

```{r, fig.height = 3, fig.width = 8}


Irr_plotDAO <- plotDAO(unitplot = expression("m"^{-1}),
                        breaks = c(0.001, 0.01, 0.1, 1, 2),
                        labels = as.character(c(0.001, 0.01, 0.1, 1, 2)),
                        runs = runs_Irr,
                        varia = "Irrigation~attenuation~")

Irr_plotDAO[[2]]

```

##### Budgets

###### C mineralization per mineralization process

```{r}

Irr_plotCbudg <- plotCbudg(runs = runs_Irr,
                          unitplot = expression("m"^{-1}),
                          varia = "Irrigation~attenuation~")

Irr_plotCbudg[[2]]

```

###### DIC production / burial inorganic

```{r}

Irr_DICinorg <- plotDICinorg(runs = runs_Irr,
                            unitplot = expression("m"^{-1}),
                            varia = "Irrigation~attenuation~")
Irr_DICinorg[2]

```

###### Alkalinity production per mineralization process - Krumins

```{r}

Irr_plotAlkProd <- plotAlkProd(runs = runs_Irr,
                            unitplot = expression("m"^{-1}),
                            varia = "Irrigation~attenuation~")
Irr_plotAlkProd[2]

```

###### Alkalinity budget -Krumins analogue

```{r}
## Alkalinity budget - Krumins analogue
Irr_plotAlkKrumins <- plot_AlkKrumins(runs = runs_Irr,
                            unitplot = expression("m"^{-1}),
                            varia = "Irrigation~attenuation~")
Irr_plotAlkKrumins[[2]]

```

###### Alkalinity budget - onlylongterm

```{r}

Irr_plotAlkLongterm <- plot_netalklongterm(runs = runs_Irr,
                                    unitplot = "(m^{-1})",
                                    varia = "Irrigation~attenuation~")
Irr_plotAlkLongterm[[2]]


ggsave("./outputfigures/irrigationalk.pdf",
       height = 4, width = 8, units = "in",
       dpi = 300)

```


#### 3.4.2.2. Second version with irrigationrate

```{r}

inputIrr <- c(0., 0.001, 0.005, 0.01, 0.05, 0.1, 0.2, 0.5, 0.75,  1)

# save each wrun
runs_Irr2 <- vector("list", length(inputIrr))

for (i in seq_along(inputIrr)) {
  
  parms_i <- baseParms
  # parms_i$irr <- inputIrr[i]
  parms_i$irr <- 0.1
  parms_i$irratt <- inputIrr[i]
  
  
  message(
    "Running Irr= ", inputIrr[i], " d^-1",
    " = ", signif(inputIrr[i], 2), " d^-1"
  )
  
  out_i <- FEMSDIA_run_steady(parms = parms_i, 
                              dynamicpH = TRUE,
                              runsteady_times = c(0, 1e7)
                              )
  
  runs_Irr2[[i]] <- list(
    Irr_cm2d = inputIrr[i],
    parms   = parms_i,
    output  = out_i
  )
}

# useful names
names(runs_Irr2) <- paste0("Irrigation_", inputIrr, "_d-1")

# save all model runs
saveRDS(runs_Irr2, file = "FEMSDIA_Irrrate_sensitivity_runs.rds")
runs_Irr2 <- readRDS("FEMSDIA_Irrrate_sensitivity_runs.rds")

```

##### Output generation

###### Plot DIC - Alkalinity - Organic C in depth for increasing reactive Fe flux

```{r, fig.height = 3, fig.width = 8}


Irr_plotDAO <- plotDAO(unitplot = expression("d"^{-1}),
                        breaks = c(0.001, 0.01, 0.1, 1, 2),
                        labels = as.character(c(0.001, 0.01, 0.1, 1, 2)),
                        runs = runs_Irr2,
                        varia = "Irrigation~rate~")

Irr_plotDAO[[2]]

```

##### Budgets

###### C mineralization per mineralization process

```{r}

Irr_plotCbudg <- plotCbudg(runs = runs_Irr2,
                          unitplot = expression("d"^{-1}),
                          varia = "Irrigation~rate~")

Irr_plotCbudg[[2]]

```

###### DIC production / burial inorganic

```{r}

Irr_DICinorg <- plotDICinorg(runs = runs_Irr2,
                            unitplot = expression("m"^{-1}),
                            varia = "Irrigation~rate~")
Irr_DICinorg[2]

```

###### Alkalinity production per mineralization process - Krumins

```{r}

Irr_plotAlkProd <- plotAlkProd(runs = runs_Irr2,
                            unitplot = expression("d"^{-1}),
                            varia = "Irrigation~rate~")
Irr_plotAlkProd[2]

```

###### Alkalinity budget -Krumins analogue

```{r}
## Alkalinity budget - Krumins analogue
Irr_plotAlkKrumins <- plot_AlkKrumins(runs = runs_Irr2,
                            unitplot = expression("d"^{-1}),
                            varia = "Irrigation~rate~")
Irr_plotAlkKrumins[[2]]

```

###### Alkalinity budget - onlylongterm

```{r}

Irr_plotAlkLongterm <- plot_netalklongterm(runs = runs_Irr2,
                                    unitplot = "(d^{-1})",
                                    varia = "Irrigation~rate~")
Irr_plotAlkLongterm[[2]]


ggsave("./outputfigures/irrigationratealk.pdf",
       height = 4, width = 8, units = "in",
       dpi = 300)

```

# 4.Figures
## 4.1. Only long term alkalinity flux, irrigation as rate

Figure 3 as used in the chapter

```{r, fig.height = 12, fig.width = 10}

#To combine

w_plotAlkLongterm <- plot_netalklongtermMAR(runs = runs_w,
                                    unitplot = expression("g m"^{-2}~"y"^{-1}),
                                    varia = "Mass~accumulation~rate~")
C_plotAlkLongterm <- plot_netalklongterm(runs = runs_Cflux,
                            unitplot = expression("(mmol m"^{-2}~"d"^{-1}~")"),
                            varia = "OC~deposition~flux~")
Fe_plotAlkLongterm <- plot_netalklongterm(runs = runs_Fe,
                                    unitplot = expression("(mmol m"^{-2}~"d"^{-1}~")"),
                                    varia = "Reactive~iron~flux~")
Ox_plotAlkLongterm <- plot_netalklongterm(runs = runs_Ox,
                                    unitplot = expression("(mmol m"^{-3}~")"),
                                    varia = "Bottomwater~oxygen~")
Biot_plotAlkLongterm <- plot_netalklongterm(runs = runs_Biot,
                                    unitplot = expression("(cm"^{2}~"y"^{-1}~")"),
                                    varia = "Biodiffusion~coeff.~")
Irr_plotAlkLongterm <- plot_netalklongterm(runs = runs_Irr2,
                                    unitplot = "(d^{-1})",
                                    varia = "Irrigation~rate~")

p1 <- Ox_plotAlkLongterm[[2]] + ggtitle("Bottomwater oxygen") + labs(tag = "a")
p2 <- w_plotAlkLongterm[[2]] + ggtitle("Mass accumulation rate") + labs(tag = "b")
p3 <- C_plotAlkLongterm[[2]] + ggtitle("OC deposition flux") + labs(tag = "c")
p4 <- Fe_plotAlkLongterm[[2]] + ggtitle("Reactive iron flux") + labs(tag = "d")
p5 <- Biot_plotAlkLongterm[[2]] + ggtitle("Biomixing rate") + labs(tag = "e")
p6 <- Irr_plotAlkLongterm[[2]] + ggtitle("Bioirrigation rate") + labs(tag = "f")

multi_panel_plot <-
  (p1 | p2) /
  (p3 | p4) /
  (p5 | p6) +
  plot_layout(guides = "collect") +
  labs(fill = "Process") +
  plot_annotation(tag_levels = "a") +
  theme(
    axis.text = element_text(size = 8),
    axis.title = element_text(size = 10),
    plot.title = element_text(hjust = 0.5, face = "bold", size = 12),
    plot.tag = element_text(face = "bold", size = 14)
  )

multi_panel_plot


ggsave("./outputfigures/longterm_irrigationrate.pdf",
       height = 12, width = 10, units = "in",
       dpi = 300)

ggsave("./outputfigures/longterm_irrigationrate.png",
       height = 12, width = 10, units = "in",
       dpi = 330)
```

# 5. Supplementary figures

## 5.1. Only long term, irrigation as attenuation

```{r, fig.height = 12, fig.width = 10}
#To combine

w_plotAlkLongterm <- plot_netalklongtermMAR(runs = runs_w,
                                    unitplot = expression("g m"^{-2}~"y"^{-1}),
                                    varia = "Mass~accumulation~rate~")
C_plotAlkLongterm <- plot_netalklongterm(runs = runs_Cflux,
                            unitplot = expression("(mmol m"^{-2}~"d"^{-1}~")"),
                            varia = "OC~deposition~flux~")
Fe_plotAlkLongterm <- plot_netalklongterm(runs = runs_Fe,
                                    unitplot = expression("(mmol m"^{-2}~"d"^{-1}~")"),
                                    varia = "Reactive~iron~flux~")
Ox_plotAlkLongterm <- plot_netalklongterm(runs = runs_Ox,
                                    unitplot = expression("(mmol m"^{-3}~")"),
                                    varia = "Bottomwater~oxygen~")
Biot_plotAlkLongterm <- plot_netalklongterm(runs = runs_Biot,
                                    unitplot = expression("(cm"^{2}~"y"^{-1}~")"),
                                    varia = "Biodiffusion~coeff.~")
Irr_plotAlkLongterm <- plot_netalklongterm(runs = runs_Irr,
                                    unitplot = "(d^{-1})",
                                    varia = "Irrigation~attenuation~")

p1 <- Ox_plotAlkLongterm[[2]] + ggtitle("Bottomwater oxygen") + labs(tag = "a")
p2 <- w_plotAlkLongterm[[2]] + ggtitle("Mass accumulation rate") + labs(tag = "b")
p3 <- C_plotAlkLongterm[[2]] + ggtitle("OC deposition flux") + labs(tag = "c")
p4 <- Fe_plotAlkLongterm[[2]] + ggtitle("Reactive iron flux") + labs(tag = "d")
p5 <- Biot_plotAlkLongterm[[2]] + ggtitle("Biomixing rate") + labs(tag = "e")
p6 <- Irr_plotAlkLongterm[[2]] + ggtitle("Bioirrigation attenuation") + labs(tag = "f")

multi_panel_plot <-
  (p1 | p2) /
  (p3 | p4) /
  (p5 | p6) +
  plot_layout(guides = "collect") +
  labs(fill = "Process") +
  plot_annotation(tag_levels = "a") +
  theme(
    axis.text = element_text(size = 8),
    axis.title = element_text(size = 10),
    plot.title = element_text(hjust = 0.5, face = "bold", size = 12),
    plot.tag = element_text(face = "bold", size = 14)
  )

multi_panel_plot


ggsave("./outputfigures/longterm_irrigationattenuation.pdf",
       height = 12, width = 10, units = "in",
       dpi = 300)

ggsave("./outputfigures/longterm_irrigationattenuation.png",
       height = 12, width = 10, units = "in",
       dpi = 330)

```

## 5.2. Mineralization budget irrigation rate

```{r, fig.height = 12, fig.width = 10}
#To combine
w_plotCbudg <- plotCbudg(runs = runs_w,
                                    unitplot = "(cm~yr^{-1})",
                                    varia = "Sediment~advection~")
C_plotCbudg <- plotCbudg(runs = runs_Cflux,
                            unitplot = expression("(mmol m"^{-2}~"d"^{-1}~")"),
                            varia = "OC~deposition~flux~")
Fe_plotCbudg <- plotCbudg(runs = runs_Fe,
                                    unitplot = expression("(mmol m"^{-2}~"d"^{-1}~")"),
                                    varia = "Reactive~iron~flux~")
Ox_plotCbudg <- plotCbudg(runs = runs_Ox,
                                    unitplot = expression("(mmol m"^{-3}~")"),
                                    varia = "Bottomwater~oxygen~")
Biot_plotCbudg <- plotCbudg(runs = runs_Biot,
                                    unitplot = "(m^{-1})",
                                    varia = "Biodiffusion~coeff.~")
Irr_plotCbudg <- plotCbudg(runs = runs_Irr2,
                          unitplot = expression("d"^{-1}),
                          varia = "Irrigation~rate~")

p1 <- w_plotCbudg[[2]] + ggtitle("Sediment advection") + labs(tag = "a")
p2 <- C_plotCbudg[[2]] + ggtitle("OC deposition flux") + labs(tag = "b")
p3 <- Fe_plotCbudg[[2]] + ggtitle("Reactive iron flux") + labs(tag = "c")
p4 <- Ox_plotCbudg[[2]] + ggtitle("Bottomwater oxygen") + labs(tag = "d")
p5 <- Biot_plotCbudg[[2]] + ggtitle("Bioturbation") + labs(tag = "e")
p6 <- Irr_plotCbudg[[2]] + ggtitle("Irrigation rate") + labs(tag = "f")

multi_panel_plot <-
  (p1 | p2) /
  (p3 | p4) /
  (p5 | p6) +
  plot_layout(guides = "collect") +
  labs(fill = "Process") +
  plot_annotation(tag_levels = "a") +
  theme(
    axis.text = element_text(size = 8),
    axis.title = element_text(size = 10),
    plot.title = element_text(hjust = 0.5, face = "bold", size = 12),
    plot.tag = element_text(face = "bold", size = 14)
  )

multi_panel_plot


ggsave("./outputfigures/mineralizationbudget_irrigationrate.pdf",
       height = 12, width = 10, units = "in",
       dpi = 300)
```

## 5.3. Krumins alkalinity fate plot

```{r, fig.height = 12, fig.width = 10}
#To combine
w_plotAlkKrumins <- plot_AlkKrumins(runs = runs_w,
                                    unitplot = "(cm~yr^{-1})",
                                    varia = "Sediment~advection~")
C_plotAlkKrumins <- plot_AlkKrumins(runs = runs_Cflux,
                            unitplot = expression("(mmol m"^{-2}~"d"^{-1}~")"),
                            varia = "OC~deposition~flux~")
Fe_plotAlkKrumins <- plot_AlkKrumins(runs = runs_Fe,
                                    unitplot = expression("(mmol m"^{-2}~"d"^{-1}~")"),
                                    varia = "Reactive~iron~flux~")
Ox_plotAlkKrumins <- plot_AlkKrumins(runs = runs_Ox,
                                    unitplot = expression("(mmol m"^{-3}~")"),
                                    varia = "Bottomwater~oxygen~")
Biot_plotAlkKrumins <- plot_AlkKrumins(runs = runs_Biot,
                                    unitplot = "(m^{-1})",
                                    varia = "Biodiffusion~coeff.~")
Irr_plotAlkKrumins <- plot_AlkKrumins(runs = runs_Irr2,
                          unitplot = expression("d"^{-1}),
                          varia = "Irrigation~rate~")

p1 <- w_plotAlkKrumins[[2]] + ggtitle("Sediment advection") + labs(tag = "a")
p2 <- C_plotAlkKrumins[[2]] + ggtitle("OC deposition flux") + labs(tag = "b")
p3 <- Fe_plotAlkKrumins[[2]] + ggtitle("Reactive iron flux") + labs(tag = "c")
p4 <- Ox_plotAlkKrumins[[2]] + ggtitle("Bottomwater oxygen") + labs(tag = "d")
p5 <- Biot_plotAlkKrumins[[2]] + ggtitle("Bioturbation") + labs(tag = "e")
p6 <- Irr_plotAlkKrumins[[2]] + ggtitle("Irrigation rate") + labs(tag = "f")

multi_panel_plot <-
  (p1 | p2) /
  (p3 | p4) /
  (p5 | p6) +
  plot_layout(guides = "collect") +
  labs(fill = "Process") +
  plot_annotation(tag_levels = "a") +
  theme(
    axis.text = element_text(size = 8),
    axis.title = element_text(size = 10),
    plot.title = element_text(hjust = 0.5, face = "bold", size = 12),
    plot.tag = element_text(face = "bold", size = 14)
  )

multi_panel_plot


ggsave("./outputfigures/kruminsalkfate_irrigationrate.pdf",
       height = 12, width = 10, units = "in",
       dpi = 300)
```