---
title: "Carbonate and alkalinity dynamics of fjord sediments - changing boundary conditions"
output: html_document
author: "Emil De Borger"
date: "2026-04-14"
---

# Introduction

This script contains simulations of alkalinity dynamics in fjords with changing
boundary conditions for chapter X in the AGU monograph "Fjord biogeochemistry" 
edited by XX.

We use the FEMSDIA model (Soetaert et al., 2026), an expansion upon the OMEXDIA
model of early diagenesis (Soetaert et al., 1996) and subsequent FESDIA model 
(Nmor et al., 2021). FEMSDIA models the dynamics of O2, C, N, P, S, Fe, and Mg in 
mineralization processes and (re)oxidation reactions, and tracks alkalinity changes
associated to said processes. A full description of the model and model code can 
be found on XX.

In this study, we set up the FEMSDIA model to simulate what we considered as average
fjord sediments, with conditions typical for (sub-)Arctic fjords. This entails generally
high porosity of 0.75, bottom water temperature of 2 degrees C, and bottom water 
conditions that generally resemble those of inflowing oceanic waters. On these 
base conditions, we test the importance of different boundary conditions that are
expected to play a strong role in governing carbonate and alkalinity dynamics and
range broadly between fjords and fjord systems globally: the sediment accumulation 
rate, the organic matter deposition rate, the reactivity of the organic matter,
the bottom water oxygen levels, the accumulation of iron, and the activities of 
fauna on the seafloor through bioturbation and bioirrigation.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(FEMSDIA)
library(ggplot2)
library(tidyr)
library(dplyr)
library(purrr)
library(ggh4x)
source("plotfuncs.R")
```

# Background settings

* Qs: CN ratio aanpassen?
*  Accumulation rates of up to 1 m y−1 have been measured in fjords where calving
  glaciers are currently retreating (Boldt, 2014).
* Difference w and LAR (Bianchi table 3 vs table 5).
* Berg et al., 2003: large amounts of S0 can be produced through reaction R13 and R14,
  and in order to keep S0 from accumulating to unrealistic levels in the anoxic sediment,
  we found it necessary to include an additional sink in the form of sulfur disproportionation
  (R21, Thamdrup and others, 1993). Sulfur disproportionating bacteria are ubiquitous 
  in marine sediments (Thamdrup and others, 1993; Canfield and Thamdrup, 1994; 
  Finster and others, 1998), and in the only direct analysis of its significance
  in marine sediments thus far, the process was found to be the main sink for S0 
  (Canfield and Thamdrup, 1996).

## NOTE SELF 22/5/2026
I'm happy with this parametrization as starting point.
Look like realistic profiles compared to publications / CANOE data.
- Anammox leave as is
- Natural sulfur is a solid
- Our model domain is 1 m, burial is expressed relative to 1 m.


```{r}

baseParms <- list(Cflux         = 4.5         ,  # Initial guess - global fjord average (https://doi.org/10.1007/s12237-025-01582-5)
                  pFast         = 0.3         ,  # Berg et al. - split of 75slow/25 fast after 37 days.
                  rFast         = 76/365.25   ,  # Soetaert et al., 1996 / was 75/365.25 in Berg but too high.
                  rSlow         = 0.1/365.25  ,  # Soetaert et al., 1996 / Berg 2003
                  NCrFdet       = 12/106      ,  # Slightly more degraded OM in mixture
                  NCrSdet       = 12/106      ,  # Slightly more degraded OM in mixture
                  NCrRdet       = 12/106      ,  # Slightly more degraded OM in mixture
                  FeOH3flux     = 0.5         ,  # Lower spectrum reported in Wehrmann et al., 2014 (DOI: 10.1016/j.gca.2014.06.007) (0.5 in Berg et al., 2003)
                  MnO2flux      = 14/365.25   ,  # Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905)
                  O2bw          = 250         , # Reasonable estimate towards well oxygenated conditions (Watts et al., 2024 https://doi.org/10.1029/2023JG007978)
                  w             = 0.005/365.25,  # Range: 0.1 - 12 cm/yr (Bianchi et al., 2020 table 3: https://doi.org/10.1016/j.earscirev.2020.103145)
                  por0          = 0.75        ,  # Depositional environment 
                  pordeep       = 0.7         ,  # Depositional environment
                  rnitri1       = 5           ,  # Own / Luna's experience that 20 (Default) is way too much.
                  rnitri2       = 5           ,  # Own / Luna's experience that 20 (Default) is way too much.
                  Twater        = 2           ,
                  biot          = 2.74e-07*12 , # based on Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905)
                  biotdepth     = 0.06        ,
                  DBL           = 0.0005      , 
                  irr           = 12/365.25   ,  # based on Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905)
                  irrdepth      = 0.003       ,
                  irratt        = 0.02        ,
                  ALKbw         = 2200        ,
                  ksO2oxic      = 3           ,
                  kinO2anox     = 1           ,
                  ksFeOH3       = 1.e+04      , # mmolFeOH3/m3 half-sat FeOH3 conc in iron reduction
                  kinFeOH3      = 1.e+04      , # mmolFeOH3/m3 half-sat FeOH3 inhibition S reduction\
                  rH2SFeox      = 1.4*10^-9 * 86400, # Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905)                  
                  ksSO4BSR      = 2e+03       , # mmolSO4/m3 half-sat SO4 conc in sulphate reduction
                  kinSO4Met     = 2e+03       , # # mmolSO4/m3 half-sat SO4 inhibition methanogenesis      
                  rH2Sox        = 4.38*10^-6  , #,# Wang and Van Capellen 1996. - > minor effect on cycles.
                  rMnox         = 0.0137      , # Van Cappellen and Wang 1996; 1.1*10^-8 * 86400 in Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905)
                  rFeox         = 0.294       ,             # Van Cappellen and Wang 1996; 1.1*10^-6 * 86400 in Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905)
                  rFeS2prod_S0  = 2.3*10^-10 * 86400 /50, # Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905) # /50 otherwise way too much FeS2
                  rFeS2prod_HS  = 8.9*10^-10 * 86400 /50, # Rickard, 1997a # /50 otherwise way too much FeS2
                  rFeMnox       = 7.5*10^-9 * 86400     , # Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905) orig: 7.5*10^-11
                  rFeSprod      = 4.7*10^-8 * 86400     , # , # Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905) -7 expo original
                  rH2SMnox      = 2*10^-9 * 86400       , # Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905)
                  rFeSox        = 6.0*10^-7 * 86400     , # Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905)
                  rFeS2ox       = 3.0*10^-11 * 86400      # Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905)
                  )

inputw <- c(0.01, 0.05, 0.1, 0.2, 0.4, 0.6, 0.8, 1, 1.5, 2, 3, 5, 10, 12) #cm/year
inputw <- inputw/100/365 # m/day


out1 <- FEMSDIA_run_steady(parms = baseParms) #, dynamicpH = TRUE)
out1budget <- FEMSDIA_get_budgetC(out1)

plot(out1, which = c("O2"), ylim = c(0.05,0))
plot(out1, which = c("DIC", "ALK"), ylim = c(0.25, 0))
plot(out1, which = c("NO3", "NH3"), ylim = c(0.025, 0))
plot(out1, which = c("MnO2", "Mn"), ylim = c(0.25, 0))
plot(out1, which = c("FeOH3", "Fe"), ylim = c(0.25, 0))
plot(out1, which = c("FeS", "FeS2"), ylim = c(0.25, 0))


phout <- FEMSDIA_get_pH(out1)
# phout
```

# The role of deposition

Investigating the effect of deposition on alkalinity dynamics is not straight-
forward, as this encompasses several aspects. Deposition is the combination 
of mineral deposits and organic matter. Mineral deposits, vary seasonally as they
are related to sources from land, in this case glacial river runoff, and snowmelt 
and precipitation on the slopes surrounding fjords. Depending on the geology of 
the fjord catchment, specific minerals relevant to biogeochemical processes, such 
as iron rich materials, are associated with this mineral deposit. In the model, 
mineral deposits contribute to the sediment advection rate "w", which affects both
solids and solutes, and the deposition fluxes of "FeOH3", and "MnO2" to a lesser 
degree.
Organic material deposition also varies, this is the result of algae blooms, both 
within and outside of the fjord, of which a fraction ends up on the seafloor. Organic 
material also comes from the land surrounding the fjords, a combination of terrestrial
plants and plant remains, but also human inputs such as waste, or fish farming 
depositions. In the model these inputs are governed by the "Cflux" parameter, the 
overall input of (reactive) organic matter, of which the "quality" - i.e. the freshness,
the reactivity, can be governed with additional parameters.
Whereas these parameters can often be strongly correlated, they do not always have
to be. As fjord environments are enormously diverse, no general rules are available
to constrain several of these parameters simultaneously. So below, we vary these
boundary conditions individually, and assess their effect on alkalinity generation
in the seafloor.

## Sediment advection rate

Measured advection rates in fjords globally range from 0.01 cm y-1 to 12 cm y-1 
(Bianchi et al., 2020 - https://doi.org/10.1016/j.earscirev.2020.103145). This range
stems from global differences, but also from within-fjord differences related to 
proximity to the glacier head. w tends to be highest closes to the head (ice-proximal)
as this region recieves input directly, of which a significant fraction sinks 
immediately.

### Runs
```{r advectionvariation}

# Vector of input w's - range from min to max reported by Bianchi et al. (2020).
inputw_cm_yr <- c(0.01, 0.05, 0.1, 0.2, 0.4, 0.6, 0.8, 1, 1.5, 2, 3, 5, 10, 12)
inputw_m_day <- inputw_cm_yr / 100 / 365.25

# save each wrun
runs_w <- vector("list", length(inputw_m_day))

for (i in seq_along(inputw_m_day)) {
  
  parms_i <- baseParms
  parms_i$w <- inputw_m_day[i]
  
  message(
    "Running w = ", inputw_cm_yr[i], " cm yr^-1",
    " = ", signif(inputw_m_day[i], 4), " m d^-1"
  )
  
  out_i <- FEMSDIA_run_steady(parms = parms_i, dynamicpH = TRUE)
  
  runs_w[[i]] <- list(
    w_cm_yr = inputw_cm_yr[i],
    w_m_day = inputw_m_day[i],
    parms   = parms_i,
    output  = out_i
  )
}

# useful names
names(runs_w) <- paste0("w_", inputw_cm_yr, "_cm_yr")
# names(runs_w) <- gsub("\\.", "p", names(runs_w))

# save all model runs
saveRDS(runs_w, file = "FEMSDIA_w_sensitivity_runs.rds")

runs_w <- readRDS("FEMSDIA_w_sensitivity_runs.rds")

```

### Output generation

#### Plot DIC - Alkalinity - Organic C in depth for increasing w

```{r, fig.height = 3, fig.width = 8}

w_plotDAO <- plotDAO(unitplot = "(cm~yr^{-1})",
                        breaks = c(0.01, 0.1, 1, 10),
                        labels = as.character(c(0.01, 0.1, 1, 10)),
                        runs = runs_w,
                        varia = "w~")

w_plotDAO[[2]]
```

#### Budgets

##### C mineralization per mineralization process

```{r}

w_plotCbudg <- plotCbudg(runs = runs_w,
                          unitplot = "(cm~yr^{-1})",
                          varia = "w~")

```

##### DIC production / burial inorganic

```{r}

w_DICinorg <- plotDICinorg(runs = runs_w,
                           unitplot = "(cm~yr^{-1})",
                           varia = "w~")
w_DICinorg[2]

```

##### Alkalinity production per mineralization process - Krumins

```{r}

w_plotAlkProd <- plotAlkProd(runs = runs_w,
                             unitplot = "(cm~yr^{-1})",
                             varia = "w~")
w_plotAlkProd[2]


```

##### Alkalinity budget -Krumins analogue

```{r}

## Alkalinity budget - Krumins analogue

w_plotAlkKrumins <- plot_AlkKrumins(runs = runs_w,
                                    unitplot = "(cm~yr^{-1})",
                                    varia = "w~")
w_plotAlkKrumins[[2]]

```

## OM deposition flux - accumulation rate

The endpoints of the range of organic matter deposition values, 0.41 - 51.2 +- 4.3
mmol C m-2 d-1 are determined by Smith et al. (2015) on the low end and Walinsky 
et al.(2009) on the high end.

```{r label, options}

# Vector of input Cflux

# Vector of input w's - range from min to max reported by Bianchi et al. (2020).
inputCmmol <- c(0.04, 0.1, 0.5, 1, 2, 5, 10, 15, 20, 25, 37.5, 50)

# save each wrun
runs_Cflux <- vector("list", length(inputCmmol))

for (i in seq_along(inputCmmol)) {
  
  parms_i <- baseParms
  parms_i$Cflux <- inputCmmol[i]
  
  message(
    "Running Cflux = ", inputCmmol[i], " mmol m^-2 yr^-1",
    " = ", signif(inputCmmol[i], 2), " mmol m^-2 yr^-1"
  )
  
  out_i <- FEMSDIA_run_steady(parms = parms_i, dynamicpH = TRUE)
  
  runs_Cflux[[i]] <- list(
    Cflux_m2day = inputCmmol[i],
    parms   = parms_i,
    output  = out_i
  )
}

# useful names
names(runs_Cflux) <- paste0("Cflux_", inputCmmol, "_mmol_/m2_/y")


# save all model runs
saveRDS(runs_Cflux, file = "FEMSDIA_Cflux_sensitivity_runs.rds")

runs_Cflux <- readRDS("FEMSDIA_Cflux_sensitivity_runs.rds")

```


### Output generation

#### Plot DIC - Alkalinity - Organic C in depth for increasing C flux

```{r, fig.height = 3, fig.width = 8}

C_plotDAO <- plotDAO(unitplot = expression("mmol m"^{-2}~"d"^{-1}),
                        breaks = c(0.01, 0.1, 1, 10),
                        labels = as.character(c(0.01, 0.1, 1, 10)),
                        runs = runs_Cflux,
                        varia = "Cflux~")

C_plotDAO[[2]]

```

#### Budgets

##### C mineralization per mineralization process

```{r}

C_plotCbudg <- plotCbudg(runs = runs_Cflux,
                          unitplot = expression("mmol m"^{-2}~"d"^{-1}),
                          varia = "Cflux~")

C_plotCbudg[[2]]

```

##### DIC production / burial inorganic

```{r}

C_DICinorg <- plotDICinorg(runs = runs_Cflux,
                            unitplot = expression("mmol m"^{-2}~"d"^{-1}),
                            varia = "Cflux~")
C_DICinorg[2]

```

##### Alkalinity production per mineralization process - Krumins

```{r}

C_plotAlkProd <- plotAlkProd(runs = runs_Cflux,
                            unitplot = expression("mmol m"^{-2}~"d"^{-1}),
                            varia = "Cflux~")
C_plotAlkProd[2]

```

##### Alkalinity budget -Krumins analogue

```{r}

## Alkalinity budget - Krumins analogue
C_plotAlkKrumins <- plot_AlkKrumins(runs = runs_Cflux,
                            unitplot = expression("mmol m"^{-2}~"d"^{-1}),
                            varia = "Cflux~")
C_plotAlkKrumins[[2]]

```

## Iron deposition - accumulation rate

Iron is deposited at different rates depending on the weathering intensity and the 
geology of the fjord catchment. Wehrmann et al. (2014 - 10.1016/j.gca.2014.06.007)
report fluxes that range from 0.44 to 29.44 mmol m-2 d-1. However, evidence shows 
that 0.6 - 12 % of this inital flux is bioavailable-reactive (Laufer-Meiser et al., 
2021 - 10.1038/s41467-021-21558-w). Over time, though, these non-reactive strongly
crystalline forms may be broken down, and become biologically available and increase
the proportion of bioavailable iron forms up to ninefold. This is not a direct depo-
sition form though, but represents an interesting source of reactive iron generate 
within the sediment itself.
Below, we test a range of reactive iron deposition fluxes of 0.026 - 3.533 mmol m-2 d-1,
corresponding to the 0.6 - 12 % range, as the our model does not include the initial
transformation of a non-reactive deposited iron fraction into bioavailable forms.


```{r label, options}

# Vector of input Fe fluxes
inputFemmol <- c(0.02, 0.05, 0.1, 0.2, 0.4, 0.6, 0.8, 1, 1.5, 2, 2.5, 3, 3.5, 4)

# Removing 4 mmol m-2 d-1, non resolved alkalinity
inputFemmol <- c(0.02, 0.05, 0.1, 0.2, 0.4, 0.6, 0.8, 1, 1.5, 2, 2.5, 3, 3.5)
 
# save each wrun
runs_Fe <- vector("list", length(inputFemmol))

for (i in seq_along(inputFemmol)) {
  
  parms_i <- baseParms
  parms_i$FeOH3flux <- inputFemmol[i]
  
  message(
    "Running Fe = ", inputFemmol[i], "mmol m^-2 yr^-1"
  )
  
  out_i <- FEMSDIA_run_steady(parms = parms_i, dynamicpH = TRUE)
  
  runs_Fe[[i]] <- list(
    Feflux_m2day = inputFemmol[i],
    parms   = parms_i,
    output  = out_i
  )
}

# useful names
names(runs_Fe) <- paste0("Fe_", inputFemmol, "_mmolm2")
# names(runs_Fe) <- gsub("\\.", "p", names(runs_Fe))

# save all model runs
saveRDS(runs_Fe, file = "FEMSDIA_Fe_sensitivity_runs.rds")

runs_Fe <- readRDS("FEMSDIA_Fe_sensitivity_runs.rds")

```

### Output generation

#### Plot DIC - Alkalinity - Organic C in depth for increasing reactive Fe flux

```{r, fig.height = 3, fig.width = 8}

Fe_plotDAO <- plotDAO(unitplot = expression("mmol m"^{-2}~"d"^{-1}),
                        breaks = c(0.01, 0.1, 1, 3),
                        labels = as.character(c(0.01, 0.1, 1, 3)),
                        runs = runs_Fe,
                        varia = "Feflux~")

Fe_plotDAO[[2]]

```

#### Budgets

##### C mineralization per mineralization process

```{r}

Fe_plotCbudg <- plotCbudg(runs = runs_Fe,
                          unitplot = expression("mmol m"^{-2}~"d"^{-1}),
                          varia = "Feflux~")

Fe_plotCbudg[[2]]

```

##### DIC production / burial inorganic

```{r}

Fe_DICinorg <- plotDICinorg(runs = runs_Fe,
                            unitplot = expression("mmol m"^{-2}~"d"^{-1}),
                            varia = "Feflux~")
Fe_DICinorg[2]

```

##### Alkalinity production per mineralization process - Krumins

```{r}

Fe_plotAlkProd <- plotAlkProd(runs = runs_Fe,
                            unitplot = expression("mmol m"^{-2}~"d"^{-1}),
                            varia = "Feflux~")
Fe_plotAlkProd[2]

```

##### Alkalinity budget -Krumins analogue

```{r}
## Alkalinity budget - Krumins analogue
Fe_plotAlkKrumins <- plot_AlkKrumins(runs = runs_Fe,
                            unitplot = expression("mmol m"^{-2}~"d"^{-1}),
                            varia = "Feflux~")
Fe_plotAlkKrumins[[2]]

```

# the role of oxygenation

Exploring the range from anoxic bottom water to fully oxygenated (default).

- Not starting at 0 mmol m-3 because that does not solve within realistic timeframe.

```{r label, options}


# Vector of input Oxygen fluxes
inputOxmmol <- c(0.1, 1, 2, 5, 10, 20, 50, 100, 150, 200, 250, 300)


# save each wrun
runs_Ox <- vector("list", length(inputOxmmol))

for (i in seq_along(inputOxmmol)) {
  
  parms_i <- baseParms
  parms_i$O2bw <- inputOxmmol[i]
  
  message(
    "Running Ox = ", inputOxmmol[i], "mmol m^-3"
  )
  
  out_i <- FEMSDIA_run_steady(parms = parms_i, dynamicpH = TRUE)
  
  runs_Ox[[i]] <- list(
    Oxbw_m2 = inputOxmmol[i],
    parms   = parms_i,
    output  = out_i
  )
}

# useful names
names(runs_Ox) <- paste0("Ox_", inputOxmmol, "_mmolm3")
# names(runs_Fe) <- gsub("\\.", "p", names(runs_Fe))

# save all model run
saveRDS(runs_Ox, file = "FEMSDIA_Ox_sensitivity_runs.rds")

runs_Ox <- readRDS("FEMSDIA_Ox_sensitivity_runs.rds")

```

### Output generation

#### Plot DIC - Alkalinity - Organic C in depth for increasing reactive Fe flux

```{r, fig.height = 3, fig.width = 8}

Ox_plotDAO <- plotDAO(unitplot = expression("mmol m"^{-3}),
                        breaks = c(0.1, 1, 30, 100, 300),
                        labels = as.character(c(0.1, 1, 30, 100, 300)),
                        runs = runs_Ox,
                        varia = "OxBW ~")

Ox_plotDAO[[2]]

```

#### Budgets

##### C mineralization per mineralization process

```{r}

Ox_plotCbudg <- plotCbudg(runs = runs_Ox,
                          unitplot = expression("mmol m"^{-2}~"d"^{-1}),
                          varia = "OxBW~")

Ox_plotCbudg[[2]]

```

##### DIC production / burial inorganic

```{r}

Ox_DICinorg <- plotDICinorg(runs = runs_Ox,
                            unitplot = expression("mmol m"^{-3}),
                            varia = "OxBW~")
Ox_DICinorg[2]

```

##### Alkalinity production per mineralization process - Krumins

```{r}

Ox_plotAlkProd <- plotAlkProd(runs = runs_Ox,
                            unitplot = expression("mmol m"^{-3}),
                            varia = "OxBW~")
Ox_plotAlkProd[2]

```

##### Alkalinity budget -Krumins analogue

```{r}
## Alkalinity budget - Krumins analogue
Ox_plotAlkKrumins <- plot_AlkKrumins(runs = runs_Ox,
                            unitplot = expression("mmol m"^{-3}),
                            varia = "OxBW~")
Ox_plotAlkKrumins[[2]]

```

# The role of faunal activities

## bioturbation

## bioirrigation

