---
title: "Carbonate and alkalinity dynamics of fjord sediments"
subtitle: "Case studies"
author: "Emil De Borger, Sebastiaan van de Velde, Ulrike Braeckman, Karline Soetaert, Astrid Hylén"
output: html_document
date: "2026-04-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(FEMSDIA)
library(ggplot2)
library(patchwork)
library(tidyr)
library(dplyr)
library(purrr)
library(ReacTran)
library(ggh4x)
library(readxl)
source("plotfuncs.R")

deepGrid2 <- setup.grid.1D(x.up = 0, dx.1 = 0.0001, N = 100, L = 2)

```

# 1. Hvalfjordur, Iceland

The carbonate-pathway, recycling-dominated case. Net benthic flux looks 
carbonate-driven, but durable/new AT is modest because much is marine-carbonate 
recycling. This is the case that teaches the two-nets distinction,  

## 1.1. Input parameters

```{r iceland}

HVparms  <-  list(Cflux         = 18.1        ,  # Initial guess - global fjord average (https://doi.org/10.1007/s12237-025-01582-5)
                  pFast         = 0.5         ,  # Berg et al. - split of 75slow/25 fast after 37 days.
                  rFast         = 50/365.25   ,  # Soetaert et al., 1996 / was 75/365.25 in Berg but too high.
                  rSlow         = 0.07/365.25 ,  # Soetaert et al., 1996 / Berg 2003
                  NCrFdet       = 12/106      ,  # Slightly more degraded OM in mixture
                  NCrSdet       = 12/106      ,  # Slightly more degraded OM in mixture
                  NCrRdet       = 12/106      ,  # Slightly more degraded OM in mixture
                  FeOH3flux     = 1.6         , 
                  MnO2flux      = 14/365      ,  # Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905)
                  O2bw          = 280         , # Reasonable estimate towards well oxygenated conditions (Watts et al., 2024 https://doi.org/10.1029/2023JG007978)
                  w             = 0.005/365.25,  # Range: 0.1 - 12 cm/yr (Bianchi et al., 2020 table 3: https://doi.org/10.1016/j.earscirev.2020.103145)
                  por0          = 0.75        ,  # Depositional environment
                  NH3bw         = 10          ,  # Bottom water ammonium
                  pordeep       = 0.7         ,  # Depositional environment
                  rnitri1       = 0.01        ,  # Own / Luna's experience that 20 (Default) is way too much.
                  rnitri2       = 0.01        ,  # Own / Luna's experience that 20 (Default) is way too much.
                  Twater        = 8.2         ,
                  biot          = 3e-8        , # based on Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905)
                  biotdepth     = 0.06        ,
                  DBL           = 0.001       , 
                  irr           = 0.2         , # based on Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905)
                  irrdepth      = 0.005       , # 0.003
                  irratt        = 0.0282      , # 0.028
                  ALKbw         = 2300        ,
                  ksO2oxic      = 6           ,
                  kinO2anox     = 2           ,
                  ksFeOH3       = 4.e+04      , # mmolFeOH3/m3 half-sat FeOH3 conc in iron reduction
                  kinFeOH3      = 1.e+04      , # mmolFeOH3/m3 half-sat FeOH3 inhibition S reduction\
                  rH2SFeox      = 1.4*10^-9 * 86400, # Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905)                  
                  ksSO4BSR      = 2e+03       , # mmolSO4/m3 half-sat SO4 conc in sulphate reduction
                  kinSO4Met     = 2e+03       , # # mmolSO4/m3 half-sat SO4 inhibition methanogenesis      
                  rH2Sox        = 4.38*10^-6  , #,# Wang and Van Capellen 1996. - > minor effect on cycles.
                  rMnox         = 0.0137      , # Van Cappellen and Wang 1996; 1.1*10^-8 * 86400 in Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905)
                  rFeox         = 0.15        ,             # Van Cappellen and Wang 1996; 0.294
                  rFeS2prod_S0  = 2.3*10^-10 * 86400 /50, # Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905) # /50 otherwise way too much FeS2
                  rFeS2prod_HS  = 8.9*10^-10 * 86400 /50, # Rickard, 1997a # /50 otherwise way too much FeS2
                  rFeMnox       = 7.5*10^-9 * 86400     , # Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905) orig: 7.5*10^-11
                  rFeSprod      = 4.7*10^-8 * 86400     ,  #4.7*10^-8 * 86400     , # , # Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905)
                  rH2SMnox      = 2*10^-9 * 86400       , # Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905)
                  rFeSox        = 6.0*10^-7 * 86400     , # Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905)
                  rFeS2ox       = 3.0*10^-11 * 86400   , # Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905)
                  CALCflux     = 1.76,
                  ARAGflux     = 0.93
                  )

HVsim <- FEMSDIA_run_steady(parms = HVparms, 
                             dynamicpH = TRUE,
                              # runsteady_times = c(0, 1e+4),
                              Grid = deepGrid2)

# saveRDS(HVsim, file = "HVsimulations.rds")

HVsim <- readRDS("HVsimulations.rds")

```


## 1.2. Assess output

### 1.2.1. Plot depth profiles with data overlay

```{r}

# plot(HVsim, ylim = c(0.2, 0),
#           which = c("DIC", "ALK", "NH3", "Fe", "FeOH3",
#                     "FeS2"))
# 
# plot(HVsim$OCpercent)

HV_profiles <- prepare_profile_comparison(
  model_output = HVsim,
  measured_file = "casedata.xlsx",
  sheet = "Profiles",
  variables = c("ALK", "DIC", "NH3"),
  measured_filters = list(
    Fjord = "Hvalfjordur",
    Station = 2
  ),
  measured_depth_col = "MidDepth",
  error_columns = c(
    ALK = "ALK_ERR",
    DIC = "DIC_ERR",
    NH3 = "NH3_ERR"
  ),
  max_depth_cm = 20
)

plot_profile_comparison(
  profile_data = HV_profiles,
  variable_labels = c(
    NH3 = expression(NH[4]^"+"),
    DIC = "DIC",
    ALK = "Alkalinity"
  ),
  max_depth_cm = 20
)

```

### 1.2.2. Plot of alkalinty generation in depth

```{r}

alk_components <- data.frame(
  depth_cm = HVsim$depth * 100,

  Ammonification =  HVsim$DINprodMin,
  Phosphate       = -HVsim$DIPprodMin,
  Denitrification =  0.8 * HVsim$Denitrific,
  FeReduction     =  8   * HVsim$Feredmin,
  MnReduction     =  4   * HVsim$MnredMin,
  SulfateReduction =       HVsim$BSRmin,
  Nitrification   = -2   * HVsim$Nitri1,
  FeOxidation     = -2   * HVsim$Feoxid,
  MnOxidation     = -2   * HVsim$Mnoxid,
  H2SOxidation    = -2   * HVsim$H2Soxid
) |>
  tidyr::pivot_longer(
    -depth_cm,
    names_to = "Process",
    values_to = "AlkRate"
  ) |>
  dplyr::filter(depth_cm <= 5)

ggplot(alk_components,
       aes(x = AlkRate, y = depth_cm, colour = Process)) +
  geom_path() +
labs(
  x = expression(
    "Alkalinity production rate" ~
      "(mmol " * A[T] ~ m^{-3} ~ d^{-1} * ")"
  ),
  y = "Depth (cm)"
) +
  xlim(-1000, 2000) +
  scale_y_reverse() +
  theme_classic()

```

### 1.2.3. Get PIC budget

```{r}

get_PIC_budget <- function(out,
                           CALCflux = NULL,
                           ARAGflux = NULL) {

  #------------------------------------------------------------
  # Integrated fluxes
  #------------------------------------------------------------

  calcite_burial   <- out$CALCdeepflux
  aragonite_burial <- out$ARAGdeepflux

  calcite_diss   <- out$TotCALCdiss
  aragonite_diss <- out$TotARAGdiss

  calcite_prod <- out$TotCALCprod

  aragonite_prod <- if ("TotARAGprod" %in% names(out))
    out$TotARAGprod
  else
    0

  #------------------------------------------------------------
  # Budget table
  #------------------------------------------------------------

  tibble(
    Mineral = c("Calcite", "Aragonite", "Total PIC"),

    Input = c(
      CALCflux,
      ARAGflux,
      CALCflux + ARAGflux
    ),

    Dissolution = c(
      calcite_diss,
      aragonite_diss,
      calcite_diss + aragonite_diss
    ),

    Production = c(
      calcite_prod,
      aragonite_prod,
      calcite_prod + aragonite_prod
    ),

    Burial = c(
      calcite_burial,
      aragonite_burial,
      calcite_burial + aragonite_burial
    )
  ) %>%

    mutate(

      Net_reaction = Production - Dissolution,

      Closure = Input +
        Production -
        Dissolution -
        Burial,

      Fraction_dissolved = Dissolution / Input,
      Fraction_produced  = Production / Input,
      Fraction_buried    = Burial / Input,

      Percent_dissolved = 100 * Fraction_dissolved,
      Percent_produced  = 100 * Fraction_produced,
      Percent_buried    = 100 * Fraction_buried

    )
}

PIC_budget_HV <- get_PIC_budget(HVsim,
                                CALCflux     = HVparms$CALCflux,
                                ARAGflux     = HVparms$ARAGflux)

PIC_budget_HV

```

### 1.2.4. Plot carbonate dynamics profiles

```{r}

get_PIC_profiles <- function(out, max_depth_cm = 25) {

  depth_cm <- out$depth * 100

  df <- tibble(
    depth_cm = depth_cm,

    CALC_diss = out$CALCdiss,
    ARAG_diss = out$ARAGdiss,
    CALC_prod = out$CALCprod,

    omegaCa = out$omegaCa,
    omegaAr = out$omegaAr,
    pH = out$pH,

    TotalCALC = out$y[,"CALC"],
    TotalARAG = out$y[,"ARAG"]
  )

  # Add aragonite production if present
  if ("ARAGprod" %in% names(out)) {
    df$ARAG_prod <- out$ARAGprod
  } else {
    df$ARAG_prod <- 0
  }

  df %>%
    filter(depth_cm <= max_depth_cm)
}

PIC_profiles_HV <- get_PIC_profiles(
  HVsim,
  max_depth_cm = 25
)

PIC_process_long <- PIC_profiles_HV %>%
  select(
    depth_cm,
    CALC_diss,
    ARAG_diss,
    CALC_prod,
    ARAG_prod
  ) %>%
  pivot_longer(
    -depth_cm,
    names_to = "Process",
    values_to = "Rate"
  )

ggplot(
  PIC_process_long,
  aes(
    x = Rate,
    y = depth_cm,
    colour = Process
  )
) +
  geom_path(linewidth = 0.9) +
  scale_y_reverse() +
  labs(
    x = "Reaction rate",
    y = "Depth (cm)",
    colour = NULL
  ) +
  theme_classic(base_size = 13)

PIC_profiles_HV %>%
  select(
    depth_cm,
    omegaCa,
    omegaAr
  ) %>%
  pivot_longer(
    -depth_cm,
    names_to = "Mineral",
    values_to = "Omega"
  ) %>%
  ggplot(
    aes(
      x = Omega,
      y = depth_cm,
      colour = Mineral
    )
  ) +
  geom_vline(
    xintercept = 1,
    linetype = "dashed",
    colour = "grey50"
  ) +
  geom_path(linewidth = 0.9) +
  scale_y_reverse() +
  labs(
    x = expression(Omega),
    y = "Depth (cm)",
    colour = NULL
  ) +
  theme_classic(base_size = 13)

ggplot(
  PIC_profiles_HV,
  aes(
    x = pH,
    y = depth_cm
  )
) +
  geom_path(linewidth = 0.9) +
  scale_y_reverse() +
  labs(
    x = "pH",
    y = "Depth (cm)"
  ) +
  theme_classic(base_size = 13)

PIC_profiles_HV %>%
  select(
    depth_cm,
    TotalCALC,
    TotalARAG
  ) %>%
  pivot_longer(
    -depth_cm,
    names_to = "Mineral",
    values_to = "Amount"
  ) %>%
  ggplot(
    aes(
      x = Amount,
      y = depth_cm,
      colour = Mineral
    )
  ) +
  geom_path(linewidth = 0.9) +
  scale_y_reverse() +
  labs(
    x = "Solid carbonate",
    y = "Depth (cm)",
    colour = NULL
  ) +
  theme_classic(base_size = 13)

```

# 2. Igaliku fjord, Greenland

I changed this from Nuup Kangerlua to Igaliku fjord, because what we want to show
is not really seen in Marius' data from Nuup. Igaliku 3 shows a relatively high
MAR and intense FeS2 formation that is desired. 

The pyrite-pathway, durable-AT case. High MAR buries FeS₂ below the bioturbated 
zone and isolates CaCO₃. 

## 2.1. Input parameters

```{r greenland}

IG3parms <- list(Cflux          = 5          , 
                  pFast         = 0.3         , 
                  rFast         = 50/365.25   , 
                  rSlow         = 0.51/365.25  ,  
                  NCrFdet       = 12/106      ,  # Slightly more degraded OM in mixture
                  NCrSdet       = 12/106      ,  # Slightly more degraded OM in mixture
                  NCrRdet       = 12/106      ,  # Slightly more degraded OM in mixture
                  FeOH3flux     = 0.5         ,  # Lower spectrum reported in Wehrmann et al., 2014 (DOI: 10.1016/j.gca.2014.06.007)
                  MnO2flux      = 14/365.25   ,  # Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905)
                  O2bw          = 280         , 
                  w             = 0.010/365.25,
                  por0          = 0.82        ,
                  pordeep       = 0.75        ,  
                  rnitri1       = 0.03        ,  
                  rnitri2       = 0.03        , 
                  Twater        = 2           ,
                  ALKbw         = 2190        ,
                  NH3bw         = 50          ,
                  biot          = 0.001/10000 , 
                  biotdepth     = 0.01        ,
                  biotatt       = 1.e-02      ,
                  DBL           = 0.001       , 
                  irr           = 0.16        ,  
                  irrdepth      = 0.02        ,
                  irratt        = 1.41/100    ,
                  NH3Ads        = 0.22        ,
                  ksO2oxic      = 5           ,
                  kinO2anox     = 2           ,
                  ksFeOH3       = 5e+04       , # mmolFeOH3/m3 half-sat FeOH3 conc in iron reduction
                  kinFeOH3      = 1e+04       , # mmolFeOH3/m3 half-sat FeOH3 inhibition S reduction\
                  rH2SFeox      = 1.4*10^-9 * 86400, # Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905)                  
                  ksSO4BSR      = 2e+03       , # mmolSO4/m3 half-sat SO4 conc in sulphate reduction
                  kinSO4Met     = 2e+03       , # # mmolSO4/m3 half-sat SO4 inhibition methanogenesis      
                  rH2Sox        = 4.38*10^-6  , #,# Wang and Van Capellen 1996. - > minor effect on cycles.
                  rMnox         = 0.0037      , # Van Cappellen and Wang 1996; 1.1*10^-8 * 86400 in Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905)
                  rFeox         = 0.1         , # Van Cappellen and Wang 1996; 1.1*10^-6 * 86400 in Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905)
                  rFeS2prod_S0  = 2.3*10^-10 * 86400 /50, # Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905) # /50 otherwise way too much FeS2
                  rFeS2prod_HS  = 8.9*10^-10 * 86400 /50, # Rickard, 1997a # /50 otherwise way too much FeS2
                  rFeMnox       = 3.5*10^-9 * 86400     , # Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905) orig: 7.5*10^-11
                  rFeSprod      = 4.2*10^-8 * 86400     ,  #4.7*10^-8 * 86400     , # , # Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905)
                  rH2SMnox      = 2.2*10^-9 * 86400     , # Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905)
                  rFeSox        = 4.0*10^-7 * 86400     , # Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905)
                  rFeS2ox       = 2.0*10^-11 * 86400 ,     # Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905)
                  CALCflux     = 0.62,
                  ARAGflux     = 0.34
                  )

IG3sim <- FEMSDIA_run_steady(parms = IG3parms, 
                              dynamicpH = TRUE,
                              # yini = IG3sim$y,
                              runsteady_times = c(0, 1e+9),
                              Grid = deepGrid2)


# test <- FEMSDIA_run_dyna(parms     = IG3parms, 
#                            times     = 0:365000,
#                            dynamicpH = TRUE)

saveRDS(IG3sim, file = "IG3simulations.rds")
IG3sim <- readRDS("IG3simulations.rds")


```

## 2.2. Assess output

### 2.2.1. Plot depth profiles with data overlay

```{r}

# plot(IG3sim, ylim = c(0.2, 0),
#           which = c("DIC", "ALK", "NH3", "Fe", "FeOH3",
#                     "FeS2"))
# 
# plot(IG3sim$OCpercent)

IG3_profiles <- prepare_profile_comparison(
  model_output = IG3sim,
  measured_file = "casedata.xlsx",
  sheet = "Profiles",
  variables = c("ALK", "DIC", "NH3"),
  measured_filters = list(
    Fjord = "Igaliku",
    Station = 3
  ),
  measured_depth_col = "MidDepth",
  error_columns = c(
    ALK = "ALK_ERR",
    DIC = "DIC_ERR",
    NH3 = "NH3_ERR"
  ),
  max_depth_cm = 20
)

plot_profile_comparison(
  profile_data = IG3_profiles,
  variable_labels = c(
    NH3 = expression(NH[4]^"+"),
    DIC = "DIC",
    ALK = "Alkalinity"
  ),
  max_depth_cm = 20
)

```

### 2.2.2. Plot of alkalinity generation in depth

```{r}
alk_components <- data.frame(
  depth_cm = IG3sim$depth * 100,

  Ammonification =  IG3sim$DINprodMin,
  Phosphate       = -IG3sim$DIPprodMin,
  Denitrification =  0.8 * IG3sim$Denitrific,
  FeReduction     =  8   * IG3sim$Feredmin,
  MnReduction     =  4   * IG3sim$MnredMin,
  SulfateReduction =       IG3sim$BSRmin,
  Nitrification   = -2   * IG3sim$Nitri1,
  FeOxidation     = -2   * IG3sim$Feoxid,
  MnOxidation     = -2   * IG3sim$Mnoxid,
  H2SOxidation    = -2   * IG3sim$H2Soxid
) |>
  tidyr::pivot_longer(
    -depth_cm,
    names_to = "Process",
    values_to = "AlkRate"
  ) |>
  dplyr::filter(depth_cm <= 20)

ggplot(alk_components,
       aes(x = AlkRate, y = depth_cm, colour = Process)) +
  geom_path() +labs(
  x = expression(
    "Alkalinity production rate" ~
      "(mmol " * A[T] ~ m^{-3} ~ d^{-1} * ")"
  ),
  y = "Depth (cm)"
) +
  scale_y_reverse() +
  theme_classic()



with(as.data.frame(IG3sim$y), {
  c(
    delta_DIC = DIC[which.max(IG3sim$depth)] - DIC[1],
    delta_ALK = ALK[which.max(IG3sim$depth)] - ALK[1],
    delta_NH3 = NH3[which.max(IG3sim$depth)] - NH3[1]
  )
})

FEMSDIA_get_budgetC(IG3sim)


```

### 2.2.3. Get PIC budget

```{r}

PIC_budget_IG <- get_PIC_budget(IG3sim,
                                CALCflux     = IG3parms$CALCflux,
                                ARAGflux     = IG3parms$ARAGflux)

PIC_budget_IG

```

### 2.2.4. Plot carbonate dynamics profiles

```{r}

PIC_profiles_IG <- get_PIC_profiles(
  IG3sim,
  max_depth_cm = 25
)

PIC_process_long <- PIC_profiles_IG %>%
  select(
    depth_cm,
    CALC_diss,
    ARAG_diss,
    CALC_prod,
    ARAG_prod
  ) %>%
  pivot_longer(
    -depth_cm,
    names_to = "Process",
    values_to = "Rate"
  )

ggplot(
  PIC_process_long,
  aes(
    x = Rate,
    y = depth_cm,
    colour = Process
  )
) +
  geom_path(linewidth = 0.9) +
  scale_y_reverse() +
  labs(
    x = "Reaction rate",
    y = "Depth (cm)",
    colour = NULL
  ) +
  theme_classic(base_size = 13)

PIC_profiles_IG %>%
  select(
    depth_cm,
    omegaCa,
    omegaAr
  ) %>%
  pivot_longer(
    -depth_cm,
    names_to = "Mineral",
    values_to = "Omega"
  ) %>%
  ggplot(
    aes(
      x = Omega,
      y = depth_cm,
      colour = Mineral
    )
  ) +
  geom_vline(
    xintercept = 1,
    linetype = "dashed",
    colour = "grey50"
  ) +
  geom_path(linewidth = 0.9) +
  scale_y_reverse() +
  labs(
    x = expression(Omega),
    y = "Depth (cm)",
    colour = NULL
  ) +
  theme_classic(base_size = 13)

ggplot(
  PIC_profiles_IG,
  aes(
    x = pH,
    y = depth_cm
  )
) +
  geom_path(linewidth = 0.9) +
  scale_y_reverse() +
  labs(
    x = "pH",
    y = "Depth (cm)"
  ) +
  theme_classic(base_size = 13)

PIC_profiles_IG %>%
  select(
    depth_cm,
    TotalCALC,
    TotalARAG
  ) %>%
  pivot_longer(
    -depth_cm,
    names_to = "Mineral",
    values_to = "Amount"
  ) %>%
  ggplot(
    aes(
      x = Amount,
      y = depth_cm,
      colour = Mineral
    )
  ) +
  geom_path(linewidth = 0.9) +
  scale_y_reverse() +
  labs(
    x = "Solid carbonate",
    y = "Depth (cm)",
    colour = NULL
  ) +
  theme_classic(base_size = 13)

```


# 3. Byfjorden, Sweden

Constructed parameters based on Watts et al., 2024 and Wittig et al., in prep.

the end-member that stress-tests the definition. No oxic zone, no fauna, FeS armouring
suppressing pyrite formation, H₂S escaping to the water column and reoxidising — 
so a naive "anoxia maximises net AT" expectation fails, and the SWI-flux-vs-durable-input
gap is at its widest. This is pedagogically the most valuable third case precisely 
because it closes the loop on 3.2 and on the non-monotonic oxygen story in 3.4.1. 

## 3.1. Input parameters

```{r sweden}

SWEparms  <- list(Cflux         = 5         ,
                  pFast         = 0.1         ,
                  rFast         = 25/365.25   ,
                  rSlow         = 0.008/365.25  ,
                  NCrFdet       = 12/106      ,  # Slightly more degraded OM in mixture
                  NCrSdet       = 12/106      ,  # Slightly more degraded OM in mixture
                  NCrRdet       = 12/106      ,  # Slightly more degraded OM in mixture
                  FeOH3flux     = 0.3         ,  # Lower spectrum reported in Wehrmann et al., 2014 (DOI: 10.1016/j.gca.2014.06.007) (0.5 in Berg et al., 2003)
                  MnO2flux      = 14/365.25   ,  # Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905)
                  O2bw          = 0.01        , # 
                  NO3bw         = 0.001       , # 0.001
                  NO2bw         = 0.001,
                  w             = 0.00076/365.25,
                  por0          = 0.86        ,
                  pordeep       = 0.78        ,
                  rnitri1       = 0.1         ,
                  rnitri2       = 0.1         ,
                  Twater        = 8.5         ,
                  biot          = 0.1e-09     , #-09
                  biotdepth     = 0.02        ,
                  biotatt       = 1.e-02      ,
                  DBL           = 0.0005      ,
                  irr           = 0.0         ,
                  irrdepth      = 0.000       ,
                  irratt        = 1/100       ,
                  ALKbw         = 4000        ,
                  DICbw         = 3950        ,
                  NH3Ads        = 0.01       ,
                  ksO2oxic      = 3           ,
                  kinO2anox     = 1           ,
                  ksFeOH3       = 1.e+04      , # mmolFeOH3/m3 half-sat FeOH3 conc in iron reduction
                  kinFeOH3      = 1.e+04      , # mmolFeOH3/m3 half-sat FeOH3 inhibition S reduction\
                  rH2SFeox      = 1.4*10^-9 * 86400, # Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905)
                  ksSO4BSR      = 2e+03       , # mmolSO4/m3 half-sat SO4 conc in sulphate reduction
                  kinSO4Met     = 2e+03       , # # mmolSO4/m3 half-sat SO4 inhibition methanogenesis
                  rH2Sox        = 4.38*10^-6  , #,# Wang and Van Capellen 1996. - > minor effect on cycles.
                  rMnox         = 0.0137      , #
                  rFeox         = 0.15        , #
                  rFeS2prod_S0  = 2.3*10^-10 * 86400 /50, # Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905) # /50 otherwise way too much FeS2
                  rFeS2prod_HS  = 8.9*10^-10 * 86400 /50, # Rickard, 1997a # /50 otherwise way too much FeS2
                  rFeMnox       = 7.5*10^-9 * 86400     , # Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905) orig: 7.5*10^-11
                  rFeSprod      = 4.7*10^-8 * 86400     ,  #4.7*10^-8 * 86400     , # , # Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905)
                  rH2SMnox      = 2*10^-9 * 86400       , # Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905)
                  rFeSox        = 6.0*10^-7 * 86400     , # Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905)
                  rFeS2ox       = 3.0*10^-11 * 86400    , # Berg et al., 2003 (DOI: 10.2475/ajs.303.10.905)
                  CALCflux     = 0.2,
                  ARAGflux     = 0.1
                  )


SWEsim <- FEMSDIA_run_steady(parms = SWEparms, 
                              dynamicpH = TRUE,
                              Grid = deepGrid2)


saveRDS(SWEsim, file = "SWEsimulations.rds")
SWEsim <- readRDS("SWEsimulations.rds")

```

## 3.2. Assess output

### 3.2.1. Plot depth profiles with data overlay

```{r}

SWE_profiles <- prepare_profile_comparison(
  model_output = SWEsim,
  measured_file = "casedata.xlsx",
  sheet = "Profiles",
  variables = c("ALK", "DIC", "NH3"),
  measured_filters = list(
    Fjord = "Byfjord",
    Station = 1
  ),
  measured_depth_col = "MidDepth",
  error_columns = c(
    ALK = "ALK_ERR",
    DIC = "DIC_ERR",
    NH3 = "NH3_ERR"
  ),
  max_depth_cm = 50
)

plot_profile_comparison(
  profile_data = SWE_profiles,
  variable_labels = c(
    NH3 = expression(NH[4]^"+"),
    DIC = "DIC",
    ALK = "Alkalinity"
  ),
  max_depth_cm = 50
)

```

### 3.2.2. Plot alkalinity generation in depth

```{r}

alk_components <- data.frame(
  depth_cm = SWEsim$depth * 100,

  Ammonification =  SWEsim$DINprodMin,
  Phosphate       = -SWEsim$DIPprodMin,
  Denitrification =  0.8 * SWEsim$Denitrific,
  FeReduction     =  8   * SWEsim$Feredmin,
  MnReduction     =  4   * SWEsim$MnredMin,
  SulfateReduction =       SWEsim$BSRmin,
  Nitrification   = -2   * SWEsim$Nitri1,
  FeOxidation     = -2   * SWEsim$Feoxid,
  MnOxidation     = -2   * SWEsim$Mnoxid,
  H2SOxidation    = -2   * SWEsim$H2Soxid
) |>
  tidyr::pivot_longer(
    -depth_cm,
    names_to = "Process",
    values_to = "AlkRate"
  ) |>
  dplyr::filter(depth_cm <= 2)

ggplot(alk_components,
       aes(x = AlkRate, y = depth_cm, colour = Process)) +
  
  geom_path()+labs(
  x = expression(
    "Alkalinity production rate" ~
      "(mmol " * A[T] ~ m^{-3} ~ d^{-1} * ")"
  ),
  y = "Depth (cm)"
) +
  scale_y_reverse() +
  theme_classic()


FEMSDIA_get_budgetC(SWEsim)


```

### 3.2.3. Get PIC budget

```{r}

PIC_budget_SWE <- get_PIC_budget(SWEsim,
                                CALCflux     = SWEparms$CALCflux,
                                ARAGflux     = SWEparms$ARAGflux)

PIC_budget_SWE

```

### 3.2.4. Plot carbonate dynamics profiles

```{r}

PIC_profiles_SWE <- get_PIC_profiles(
  SWEsim,
  max_depth_cm = 25
)

PIC_process_long <- PIC_profiles_SWE %>%
  select(
    depth_cm,
    CALC_diss,
    ARAG_diss,
    CALC_prod,
    ARAG_prod
  ) %>%
  pivot_longer(
    -depth_cm,
    names_to = "Process",
    values_to = "Rate"
  )

ggplot(
  PIC_process_long,
  aes(
    x = Rate,
    y = depth_cm,
    colour = Process
  )
) +
  geom_path(linewidth = 0.9) +
  scale_y_reverse() +
  labs(
    x = "Reaction rate",
    y = "Depth (cm)",
    colour = NULL
  ) +
  theme_classic(base_size = 13)

PIC_profiles_SWE %>%
  select(
    depth_cm,
    omegaCa,
    omegaAr
  ) %>%
  pivot_longer(
    -depth_cm,
    names_to = "Mineral",
    values_to = "Omega"
  ) %>%
  ggplot(
    aes(
      x = Omega,
      y = depth_cm,
      colour = Mineral
    )
  ) +
  geom_vline(
    xintercept = 1,
    linetype = "dashed",
    colour = "grey50"
  ) +
  geom_path(linewidth = 0.9) +
  scale_y_reverse() +
  labs(
    x = expression(Omega),
    y = "Depth (cm)",
    colour = NULL
  ) +
  theme_classic(base_size = 13)

ggplot(
  PIC_profiles_SWE,
  aes(
    x = pH,
    y = depth_cm
  )
) +
  geom_path(linewidth = 0.9) +
  scale_y_reverse() +
  labs(
    x = "pH",
    y = "Depth (cm)"
  ) +
  theme_classic(base_size = 13)

PIC_profiles_SWE %>%
  select(
    depth_cm,
    TotalCALC,
    TotalARAG
  ) %>%
  pivot_longer(
    -depth_cm,
    names_to = "Mineral",
    values_to = "Amount"
  ) %>%
  ggplot(
    aes(
      x = Amount,
      y = depth_cm,
      colour = Mineral
    )
  ) +
  geom_path(linewidth = 0.9) +
  scale_y_reverse() +
  labs(
    x = "Solid carbonate",
    y = "Depth (cm)",
    colour = NULL
  ) +
  theme_classic(base_size = 13)

```

# 4. Synthesis figure

## 4.1. Extract metrics needed for plot

```{r synthesis}

library(dplyr)
library(tidyr)
library(ggplot2)
library(viridis)

get_synthesis_metrics <- function(out) {

  # mmol CaCO3 m-2 d-1:
  # positive = net carbonate dissolution
  carbonate_balance <-
    out$TotCALCdiss +
    out$TotARAGdiss -
    out$TotCALCprod

  # mmol FeS2 m-2 d-1
  pyrite_burial <- out$FeS2deepflux

  # Combined durable alkalinity effect
  durable_AT <-
    2 * carbonate_balance +
    4 * pyrite_burial
  
  # At flux from model
  ATflux <- out$ALKflux

  tibble(
    Carbonate_balance = as.numeric(carbonate_balance),
    Pyrite_burial = as.numeric(pyrite_burial),
    Durable_AT = as.numeric(durable_AT),
    Model_AT = -as.numeric(ATflux)
  )
}

fjord_points <- bind_rows(


  get_synthesis_metrics(HVsim) |>
    mutate(Fjord = "Hvalfjörður"),

  get_synthesis_metrics(IG3sim) |>
    mutate(Fjord = "Igaliku Fjord"),
  
  get_synthesis_metrics(SWEsim) |>
    mutate(Fjord = "Byfjord")
)

fjord_points
fjord_points$Fjord <- factor(fjord_points$Fjord, levels = c("Hvalfjörður", "Igaliku Fjord", "Byfjord"))
```

## 4.2. A panel - plot with filled background

```{r label, options}

x_range <- range(
  fjord_points$Carbonate_balance,
  na.rm = TRUE
)

y_range <- range(
  fjord_points$Pyrite_burial,
  na.rm = TRUE
)

x_buffer <- max(diff(x_range) * 0.25, 0.05)
y_buffer <- max(diff(y_range) * 0.25, 0.01)

carbonate_vector <- seq(
  from = -1,
  to = 3.3,
  length.out = 200
)

pyrite_vector <- seq(
  from = 0,
  to = 1,
  length.out = 200
)

background_df <- tidyr::expand_grid(
  Carbonate_balance = carbonate_vector,
  Pyrite_burial = pyrite_vector
) %>%
  dplyr::mutate(
    Durable_AT =
      2 * Carbonate_balance +
      4 * Pyrite_burial
  )

```

```{r, fig.height=5, fig.width=10}


synthesis_plot <- ggplot() +

  geom_raster(
    data = background_df,
    aes(
      x = Carbonate_balance,
      y = Pyrite_burial,
      fill = Durable_AT
    ),
    interpolate = TRUE
  ) +

  geom_contour(
    data = background_df,
    aes(
      x = Carbonate_balance,
      y = Pyrite_burial,
      z = Durable_AT
    ),
    colour = "black",
    linewidth = 0.45,
    alpha = 0.3
  ) +

geom_point(
  data = fjord_points,
  aes(
    x = Carbonate_balance,
    y = Pyrite_burial,
    shape = Fjord
  ),
  size = 4.8,
  fill = "lightgrey",
  colour = "black",
  stroke = 1.5
) +
scale_shape_manual(
  values = c(
    "Hvalfjörður" = 24,
    "Igaliku Fjord" = 22,
    "Byfjord" = 21
  )
) +

  geom_text(
    data = fjord_points,
    aes(
      x = Carbonate_balance,
      y = Pyrite_burial,
      label = Fjord
    ),
    nudge_x = 0.03 * diff(range(carbonate_vector)),
    nudge_y = 0.03 * diff(range(pyrite_vector)),
    hjust = 0,
    size = 4
  ) +
  scale_x_continuous(
  limits = c(-1, 3.3),
  breaks = c(-1, 0, 1, 2, 3)
) +

  scale_fill_viridis_c(
  option = "D",
  limits = range(background_df$Durable_AT),
  breaks = c(-2, 0, 2, 4, 6, 8, 10),
  name = expression("Net benthic " * A[T] ~ "flux"~~~"(mmol m"^{-2}~d^{-1}*")")
) +

  labs(
  x = expression(
    "Net CaCO"[3]~"dissolution - precipitation"~
      "(mmol CaCO"[3]~"m"^{-2}~d^{-1}*")"
  ),

    y = expression(
      "FeS"[2]~" burial (mmol m"^{-2}~d^{-1}*")"
    )
  ) +

  coord_cartesian(
    expand = FALSE,
    clip = "off"
  ) +

  theme_classic(base_size = 13) +

  theme(
    axis.text = element_text(colour = "black"),
    legend.position = "right",
    plot.margin = margin(10, 50, 10, 10)
  )

synthesis_plot

```

## 4.3. B panel - plot of comparison net and true AT flux

```{r label, options}


ggplot(fjord_points,
       aes(x = Durable_AT,
           y = Model_AT,
           shape = Fjord)) +
  geom_abline(slope = 1, intercept = 0,
              linetype = "dashed", colour = "grey50") +
  geom_point(size = 4.5, fill = "white", colour = "black", stroke = 1.3) +
  geom_text(aes(label = Fjord),
            nudge_x = 0.15,
            show.legend = FALSE) +
  coord_equal() +
  labs(
    x = expression("Net benthic " * A[T] ~ "flux"~~~"(mmol m"^{-2}~d^{-1}*")"),
    y = expression("Full benthic " * A[T] ~ "flux"~~~"(mmol m"^{-2}~d^{-1}*")")
  ) +
  theme_bw(base_size = 14)

```

## 4.4. Combining both plots

```{r , fig.height= 6, fig.width=10}

library(ggplot2)
library(patchwork)

# ------------------------------------------------------------
# Panel A
# ------------------------------------------------------------

synthesis_plot <- synthesis_plot +

  guides(
    shape = guide_legend(
      title.position = "top",
      title.hjust = 0.5,
      nrow = 1,
      order = 1,
      override.aes = list(
        size = 4,
        fill = "lightgrey"
      )
    ),
    fill = guide_colourbar(
      title.position = "top",
      title.hjust = 0.5,
      barwidth = grid::unit(8, "cm"),
      barheight = grid::unit(0.45, "cm"),
      direction = "horizontal",
      order = 2
    )
  ) +

  theme(
    legend.position = "bottom",
    legend.box = "vertical",
    legend.box.just = "center",
    plot.margin = margin(10, 10, 5, 10)
  )


# ------------------------------------------------------------
# Panel B
# ------------------------------------------------------------

comparison_plot <- ggplot(
  fjord_points,
  aes(
    x = Durable_AT,
    y = Model_AT,
    shape = Fjord
  )
) +

  geom_abline(
    slope = 1,
    intercept = 0,
    linetype = "dashed",
    colour = "grey55",
    linewidth = 0.5
  ) +

  geom_point(
    size = 4.3,
    fill = "lightgrey",
    colour = "black",
    stroke = 1.3
  ) +

  scale_shape_manual(
    values = c(
      "Byfjord" = 21,
      "Hvalfjörður" = 24,
      "Igaliku Fjord" = 22
    )
  ) +

  coord_equal(
    xlim = c(-0.2, 11),
    ylim = c(0, 11),
    expand = FALSE
  ) +

  scale_x_continuous(
    breaks = c(0, 2.5, 5, 7.5, 10)
  ) +

  scale_y_continuous(
    breaks = c(0, 2.5, 5, 7.5, 10)
  ) +

  labs(
    y = expression(
      "Net benthic " * A[T] ~ "flux" ~
        "(mmol m"^{-2} ~ d^{-1} * ")"
    ),
    x = expression(
      "Net " * A[T] ~ "input to the ocean" ~
        "(mmol m"^{-2} ~ d^{-1} * ")"
    )
  ) +

  theme_classic(base_size = 12) +

  theme(
    legend.position = "none",
    axis.text = element_text(colour = "black"),
    plot.margin = margin(10, 10, 5, 10)
  )

comparison_plot


library(patchwork)
library(grid)

# Remove the duplicate legend from panel B
comparison_plot <- comparison_plot +
  guides(shape = "none") +
  theme(
    legend.position = "none"
  )

# Format the two legends from panel A
synthesis_plot <- synthesis_plot +
  guides(shape = guide_legend(
      title.position = "top",
      title.hjust = 0.5,
      nrow = 1,
      byrow = TRUE,
      order = 1,
      override.aes = list(
        size = 4,
        fill = "lightgrey"
      )
    ),
    fill = guide_colourbar(
      title.position = "top",
      title.hjust = 0.5,
      direction = "horizontal",
      barwidth = unit(7, "cm"),
      barheight = unit(0.4, "cm"),
      order = 2
    )
  ) +
  theme(
    legend.position = "bottom",
    legend.box = "vertical",
    legend.box.just = "center",
    legend.justification = "center"
  )

# Combine
combined_plot_horizontallegend <- (
    synthesis_plot +
    comparison_plot +        
    plot_layout(
      widths = c(1.65, 1),
      guides = "collect"
    )
) +
  plot_annotation(
    tag_levels = "a"
  ) &
  theme(
    legend.position = "bottom",
    legend.box = "horizontal",
    legend.box.just = "center",
    legend.justification = "center",
    plot.tag = element_text(
      face = "bold",
      size = 14
    )
  )

combined_plot_horizontallegend

ggsave("./outputfigures/combined_plot_horizontallegend.png",
       height = 6, width = 10, units = "in",
       dpi = 330)


```